---
title: "Prepare figure 1"
author: "Stefano Giulieri"
date: "15/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(paste0(here::here(), "/6-InToxSa_paper/Figure1_prepare"))
getwd()
```

Here we generate proof of principle figure for the paper with model PI uptake for:

1. JE2
2. NE1532 (agrA JE2)
3. Non infected cells

```{r}
library(tidyverse)
library(ggridges)
library(ggh4x)
library(skimr)
library(corrr)
library(psych)
library(GGally)
rm(list = ls())

source("../../0-Functions/all_functions.R")
```

# Import PI data

```{r}
PI_data <- readRDS("../../5-Data_processing/processed_data/allelic_exchange/PI_kinetics/PI_kinetics_plateGP-C1.Rda")
```

# Plot data of interest

```{r}
colorscale <- c(JE2 = "#045a8d", NE1532 = "#b30000", `Non infected` = "#009900")

df_plot <- PI_data %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  filter(timepoint < 60*20)

plot_cell_death(df_plot, fitted = T)

# Parameters for vline
PI_parameters <- readRDS("../../5-Data_processing/processed_data/allelic_exchange/PI_parameters/PI_sample_parameters_plateGP-C1.Rda")
# slope (b)
b <- PI_parameters$max_rate_death_mean[which(PI_parameters$sample_id == "JE2")]*60

# intercept (a): a = y - bx

x <- PI_parameters$time_of_max_rate_death_mean[which(PI_parameters$sample_id == "JE2")]/60
# Alternative way
# Plot data
PI_data %>%
  filter(sample_id == "JE2") %>%
  group_by(timepoint) %>%
  summarise(fitted.deriv1 = mean(fitted.deriv1)) %>%
  ggplot(aes(x = timepoint/60, y= fitted.deriv1)) +
  geom_point() +
  geom_vline(aes(xintercept = 1/60*timepoint[which(fitted.deriv1 == max(fitted.deriv1))])) +
   geom_hline(aes(yintercept = max(fitted.deriv1))) +
  labs(x = "Time (hour)", y = "PI uptake increase (Death rate)",
       title = "JE2: change in rate of PI uptake during the experiment",
       subtitle = "First derivative from local fitting") +
  theme_bw()
# Extract data
x <- PI_data %>%
  filter(sample_id == "JE2") %>%
  group_by(timepoint) %>%
  summarise(fitted.deriv1 = mean(fitted.deriv1)) %>%
  ungroup() %>%
  filter(fitted.deriv1 == max(fitted.deriv1)) %>%
  .$timepoint / 60
y <- PI_data %>%
  filter(sample_id == "JE2") %>%
  group_by(timepoint) %>%
  summarise(f_signal.fitted = mean(f_signal.fitted),
            fitted.deriv1 = mean(fitted.deriv1)) %>%
  ungroup() %>%
  filter(fitted.deriv1 == max(fitted.deriv1)) %>%
  .$f_signal.fitted 
a <- y - b*x




p <- df_plot %>%
      ggplot(aes(x = timepoint/60, y = f_signal.fitted, fill = sample_id, colour = sample_id)) +
      #stat_summary(geom = "ribbon", fun.data = mean_cl_normal, fun.args=list(conf.int=0.95), fill = "#045a8d", alpha = 0.2) + # Manually checked and give same results, but 95% CI appear to be too narrow 
      stat_summary(geom = 'ribbon', fun.data = 'mean_sdl', fun.args = list(mult = 1),  alpha = 0.2) + # Use standard deviation instead
      stat_summary(geom = "point", fun = "mean", size = 1) +
  geom_vline(xintercept = 5, linetype = "dashed", alpha = .5) +
  geom_vline(xintercept = 20, linetype = "dashed", alpha = .5) +
    geom_abline(intercept = a, slope = b) +
  scale_fill_manual(values = colorscale) +
  scale_color_manual(values = colorscale) +
      labs(title = "",
           subtitle = "",
           x = "Time (hour)",
           y = "PI uptake") +
      theme_bw() +
  theme(legend.position = "none",
        text = element_text(face = "bold"),
        panel.grid = element_blank())
p


# ggsave("figures/PI_prototype_curve.pdf", width = 2.6, height = 2.4)
```

# Try to find the best plot

```{r}
exp <- unique(PI_data$experiment_id)

plot_list <- list()

for (e in exp){
  df <- df_plot %>%
  filter(experiment_id == e)

# Parameters for vline
# slope (b)
b <- PI_data %>%
  filter(sample_id == "JE2") %>%
  filter(experiment_id == e) %>%
  filter(fitted.deriv1 == max(fitted.deriv1)) %>%
  .$fitted.deriv1 * 60

# intercept (a): a = y - bx
x <- PI_data %>%
  filter(sample_id == "JE2") %>%
  filter(experiment_id == e) %>%
  group_by(timepoint) %>%
  summarise(fitted.deriv1 = mean(fitted.deriv1)) %>%
  ungroup() %>%
  filter(fitted.deriv1 == max(fitted.deriv1)) %>%
  .$timepoint / 60
y <- PI_data %>%
  filter(sample_id == "JE2") %>%
  filter(experiment_id == e) %>%
  group_by(timepoint) %>%
  summarise(f_signal.fitted = mean(f_signal.fitted),
            fitted.deriv1 = mean(fitted.deriv1)) %>%
  ungroup() %>%
  filter(fitted.deriv1 == max(fitted.deriv1)) %>%
  .$f_signal.fitted 
a <- y - b*x

p <- df %>%
      ggplot(aes(x = timepoint/60, y = f_signal.fitted, fill = sample_id, colour = sample_id)) +
      stat_summary(geom = 'ribbon', fun.data = 'mean_sdl', fun.args = list(mult = 1),  alpha = 0.2) + # Use standard deviation instead
      stat_summary(geom = "point", fun = "mean", size = 1) +
  geom_vline(xintercept = 5, linetype = "dashed", alpha = .5) +
  geom_vline(xintercept = 20, linetype = "dashed", alpha = .5) +
    geom_abline(intercept = a, slope = b) +
    stat_summary(geom = "area", fun = "mean", size = 1, fill = "#969696", alpha = .1, linetype = "blank") +
  scale_fill_manual(values = colorscale) +
  scale_color_manual(values = colorscale) +
      labs(x = "Time (hour)",
           y = "PI uptake",
           title = e) +
      theme_bw() +
  theme(legend.position = "none",
        text = element_text(face = "bold"),
        panel.grid = element_blank())

plot_list[[e]] <- p

}

plot_list %>%
  patchwork::wrap_plots()
```

The last plot (`exp220328_GP-C1`) is probably the most representative

```{r}
p <- plot_list[[5]]
p

# ggsave("figures/PI_prototype_curve.pdf", width = 2.6, height = 2.4)
```

How many replicates

```{r}
p$data %>%
  group_by(sample_id) %>%
  summarise(n_replicates = n_distinct(well_id))
```

Plot variation across replicates

AUC death only

```{r}
PI_parameters_with_replicates <- readRDS("../../Mutants_2022/processed_data/PI_parameters/PI_parameters_plateGP-C1.Rda")

PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = sample_id, y = AUC_death)) +
  geom_violin(aes(fill = sample_id), alpha =.5) +
  geom_boxplot(outlier.shape = NA, width = .1, alpha = .5, fill = "white") +
   scale_fill_manual(values = colorscale) +
  theme_bw()

PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = sample_id, y = AUC_death)) +
  geom_violin(aes(fill = sample_id), alpha =.5) +
  geom_boxplot(outlier.shape = NA, width = .1, alpha = .5, fill = "white") +
  facet_wrap(~experiment_id) +
   scale_fill_manual(values = colorscale) +
  theme_bw()

PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = sample_id, y = AUC_death)) +
  geom_violin( alpha =.5) +
  geom_boxplot(aes(fill = experiment_id),  width = .5, alpha = .5) +
  scale_fill_brewer() +
  theme_bw()

# density plots
PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = AUC_death)) +
 geom_density(aes(fill = sample_id), alpha = .5) +
   scale_fill_manual(values = colorscale) +
  theme_bw()

PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = AUC_death, fill = experiment_id)) +
 geom_density(alpha = .5) +
  facet_wrap(~sample_id, scales = "free_y") +
   scale_fill_brewer() +
  theme_bw()

PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = AUC_death, y = sample_id, fill = sample_id)) +
geom_density_ridges(alpha = .5) +
   scale_fill_manual(values = colorscale) +
  theme_bw()

PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = AUC_death, y = sample_id, fill = sample_id)) +
geom_density_ridges(alpha = .5) +
  facet_wrap(~experiment_id) +
 scale_fill_manual(values = colorscale) +
  theme_bw()

PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  ggplot(aes(x = AUC_death, y = sample_id, fill = experiment_id)) +
geom_density_ridges(alpha = .5) +
   scale_fill_brewer() +
  theme_bw()
```

Now plot multiple parameters

```{r}
df_plot <- PI_parameters_with_replicates %>%
  filter(sample_id %in% c("JE2", "NE1532", "Non infected")) %>%
  select(sample_id, experiment_id,  AUC_death, max_rate_death, max_death, time_of_max_death, min_death, time_of_min_death) %>%
  pivot_longer(-c(sample_id, experiment_id), names_to = "parameter", values_to = "value") 

df_plot %>%
  ggplot(aes(x = sample_id, y = value)) +
  geom_violin(aes(fill = sample_id), alpha =.5) +
  geom_boxplot(outlier.shape = NA, width = .1, alpha = .5, fill = "white") +
  facet_wrap(~parameter, scales = "free_y") +
   scale_fill_manual(values = colorscale) +
  theme_bw()

df_plot %>%
  ggplot(aes(x = sample_id, y = value)) +
  geom_violin(aes(fill = sample_id), alpha =.5) +
  geom_boxplot(outlier.shape = NA, width = .1, alpha = .5, fill = "white") +
  facet_grid(parameter~experiment_id, scales = "free_y") +
   scale_fill_manual(values = colorscale) +
  theme_bw()

df_plot %>%
  ggplot(aes(x = sample_id, y = value)) +
  geom_violin( alpha =.5) +
  geom_boxplot(aes(fill = experiment_id),  width = .5, alpha = .5) +
  scale_fill_brewer(guide = "none") +
  facet_wrap(~parameter, scales = "free_y") +
  theme_bw()
```

Density plots

```{r}
df_plot %>%
 ggplot(aes(x = value)) +
 geom_density(aes(fill = sample_id), alpha = .5) +
  facet_wrap(~parameter, scales = "free") +
   scale_fill_manual(values = colorscale) +
  theme_bw()

df_plot %>%
  ggplot(aes(x = value, fill = experiment_id)) +
 geom_density(alpha = .5) +
  facet_wrap(parameter~sample_id, scales = "free") +
   scale_fill_brewer() +
  theme_bw()

df_plot %>%
 ggplot(aes(x = value, y = sample_id, fill = sample_id)) +
geom_density_ridges(alpha = .5) +
  facet_wrap(~parameter, scales = "free") +
   scale_fill_manual(values = colorscale) +
  theme_bw()

df_plot %>%
  ggplot(aes(x = value, y = sample_id, fill = experiment_id)) +
geom_density_ridges(alpha = .5) +
  facet_wrap(~parameter, scales = "free_x") +
   scale_fill_brewer(guide = "none") +
  theme_bw()
```

Export plots

Figure 1 panel B

```{r}
p <- df_plot %>%
  filter(parameter %in% c("AUC_death", "max_death", "max_rate_death")) %>%
  mutate(sample_id = fct_relevel(sample_id, c("JE2", "NE1532"))) %>%
  ggplot(aes(x = sample_id, y = value)) +
  geom_violin(aes(colour = sample_id), alpha =.5, size = 1, width = 1.3) +
  geom_boxplot(aes(fill = experiment_id),  width = .5, alpha = .5) +
  scale_x_discrete(labels = c("JE2", "agrA\nmutant", "Non\ninfected")) +
  scale_fill_brewer(guide = "none") +
    scale_colour_manual(values = colorscale, guide = "none") +
  facet_wrap(~parameter, scales = "free_y",
             labeller = labeller(parameter = c(AUC_death = "AUC", max_death = "Peak", max_rate_death = "Maximum rate"))) +
  # facetted_pos_scales(y = list(
  #   parameter == "max_rate_death" ~ scale_y_continuous(limits = c(0,.01)))) +
  labs(x = "", y = "") +
  theme_bw()
p

ggsave("figures/parameters_variation_plate-GP-C1_3cols.pdf")

p +
  facet_wrap(~parameter, scales = "free_x", ncol = 1,
             labeller = labeller(parameter = c(AUC_death = "AUC", max_death = "Peak", max_rate_death = "Maximum rate"))) +
  coord_flip() +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"))

ggsave("figures/parameters_variation_plate-GP-C1_1col.pdf", height = 8, width = 6)
ggsave("figures/parameters_variation_plate-GP-C1_1col.png", height = 8, width = 6)

p <- df_plot %>%
  filter(parameter %in% c("AUC_death", "max_death")) %>%
  mutate(sample_id = fct_relevel(sample_id, c("JE2", "NE1532"))) %>%
  ggplot(aes(x = sample_id, y = value)) +
  geom_violin(aes(colour = sample_id), alpha =.5, size = 1, width = 1.3) +
  geom_boxplot(aes(fill = experiment_id),  width = .5, alpha = .5) +
  scale_x_discrete(labels = c("JE2", "agrA\nmutant", "Non\ninfected")) +
  scale_fill_brewer(guide = "none") +
    scale_colour_manual(values = colorscale, guide = "none") +
  facet_wrap(~parameter, scales = "free_y",
             labeller = labeller(parameter = c(AUC_death = "AUC", max_death = "Peak", max_rate_death = "Maximum rate"))) +
  # facetted_pos_scales(y = list(
  #   parameter == "max_rate_death" ~ scale_y_continuous(limits = c(0,.01)))) +
  labs(x = "", y = "") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"))
p

# ggsave("figures/parameters_variation_plate-GP-C1_2param_2cols.pdf", height = 5, width = 6)
# ggsave("figures/parameters_variation_plate-GP-C1_2params_2cols.png", height = 5, width = 6)
```

All parameters

```{r}
p <- df_plot %>%
  mutate(sample_id = fct_relevel(sample_id, c("JE2", "NE1532"))) %>%
  ggplot(aes(x = sample_id, y = value)) +
  geom_violin(aes(colour = sample_id), alpha =.5, size = 1, width = 1.3) +
  geom_boxplot(aes(fill = experiment_id),  width = .5, alpha = .5) +
  scale_x_discrete(labels = c("JE2", "agrA\nmutant", "Non\ninfected")) +
  scale_fill_brewer(guide = "none") +
    scale_colour_manual(values = colorscale, guide = "none") +
  facet_wrap(~parameter, scales = "free_y",
             labeller = labeller(parameter = c(AUC_death = "AUC", max_death = "Peak", max_rate_death = "Maximum rate",
                                               min_death = "Trough", time_of_max_death = "Time of peak", time_of_min_death = "Time of trough"))) +
  labs(x = "", y = "") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"))
p

# ggsave("figures/parameters_variation_plate-GP-C1_all_params.pdf", height = 10, width = 12)
# ggsave("figures/parameters_variation_plate-GP-C1_all_params.png", height = 10, width = 12)
```

# Stats

Basic descriptive statistics

```{r}
df_stats <- PI_parameters_with_replicates %>%
  select(sample_id, ends_with("death")) %>%
  filter(sample_id %in% c("JE2", "TOX-4", "Non infected")) %>%
  as_tibble()

df_stats %>%
  group_by(sample_id) %>%
  skim() %>%
  as_tibble() %>%
  mutate(coeff_variation = numeric.sd/numeric.mean) %>%
  select(sample_id, skim_variable, numeric.mean, numeric.sd, coeff_variation) %>%
  filter(skim_variable %in% c("AUC_death", "max_death"))
```

Correlation between replicates, Crobach's alpha etc

See tutorial here https://randilgarcia.github.io/datic-intro-to-r/cor_reliability_ttest.html#cronbach%E2%80%99s_alpha

```{r}
# AUC for JE2
df <- PI_parameters_with_replicates %>%
  filter(sample_id == "JE2") %>%
  select(well_id, AUC_death, experiment_id) %>%
  group_by(experiment_id) %>%
  mutate(replicate = row_number()) %>%
  select(-well_id) %>%
  pivot_wider(names_from = "experiment_id", values_from = "AUC_death")

alpha(df)
```

Correlation of *fitted* data

```{r}
df <- PI_data %>%
  filter(sample_id == "JE2") %>%
  ungroup() %>%
  select(well_id, timepoint, experiment_id, f_signal.fitted, plate_replicate)

df_wide <- df %>%
  select(timepoint, well_id, f_signal.fitted) %>%
  pivot_wider(names_from = "well_id", values_from = f_signal.fitted)

alpha(df_wide %>% select(-timepoint))

df_wide %>%
  select(-timepoint) %>%
  ggpairs()
```

By experiment

```{r}
df_wide <- df %>%
  group_by(timepoint, experiment_id) %>%
  summarise(f_signal.fitted = mean(f_signal.fitted, na.rm = T)) %>%
  pivot_wider(names_from = "experiment_id", values_from = f_signal.fitted)

alpha(df_wide %>% select(-timepoint))

df_wide %>%
  ggpairs()
```

Correlation is difficult to interpret for the fitted curves. We could calculate correlations for the entire dataset if we wanted.
