---
title: "Concatenate and check cell death data"
author: "Stefano Giulieri"
date: "05/11/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This is a script to concatenate all cell death dataframes in a single file and then perform data checks. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
msg <- stringr::str_c("My directory is ", here::here())
message(msg)
```

```{r message=FALSE}
library(tidyverse)
library(magrittr)
rm(list = ls())
source("Functions/all_functions.R")
```

# Concatenate single parsed dataframes 

As of 24 February 2020, we have now parsed cell death data as single experiments that can now be concatenated.
We first generate a list of experiments to retain: basically all with 40k cells and after excluding problematic experiments.

## List of available parsed cell death dataframes

```{r}
c <- list.files("Data_parsing/parsed_experiments", recursive = T, full.names = TRUE) %>% str_subset("PI")

c_files <- file.info(c) %>%
  rownames_to_column("file") %>%
  as_tibble() %>%
  transmute(fname = basename(file), mtime, ctime, atime)

knitr::kable(c_files)
```

We have a total of `r length(c)` experiments with PI data.

## Concatenate raw data

```{r message=FALSE}
# all cell death data
all_c_data_raw <- purrr::map(c, function(c){
  df <- read_csv(c) %>%
    mutate(plate = as.character(plate))
}) %>%
  bind_rows()
# experiments with less than 40k cells
cells_low <- all_c_data_raw %>%
  filter(cell_number != 4e4) %>%
  select(plate, experiment) %>%
  distinct()
cells_low
# experiments excluded for other reasons: exp200228 (first experiment for plate M1)
all_c_data <- all_c_data_raw %>%
  filter(!experiment %in% c(cells_low$experiment, "exp200228"))
```

We excluded 5 experiments: four with less than 40k cells and one that failed.

## Check concatenated file

```{r}
# included plates / experiments
all_c_data %>%
  select(plate, experiment) %>%
  distinct() %>%
  knitr::kable(row.names = T)
```

Three experiments (9 replicates) for each of plates GP4 and GP7. Two experiments (6 replicates) for each of plates 1-6, P2 and GP1-GP3 and GP5-GP6 and one experiment (one replicate) for each of plates P1, M1, M2. In total 33 experiments.

```{r}
# Number of wells per experiment
all_c_data %>%
  group_by(plate, experiment) %>%
  summarise(n_wells = n_distinct(well)) %>%
  knitr::kable()
```

We have now confirmed that the concatenated dataframe includes all experiments of interest and the all wells have been parsed.

```{r}
saveRDS(all_c_data,
          "Data_concatenation/dataframes/concatenated_checked/all_cell_death_data_parsed.Rda")
```

## Remove outliers manually

Outliers were identified manually in plates 1-6 and plates M1-M2. No outliers were identified in plates P1 and P2.
Here, we create a complete dataframe of manually identified outliers

This part of the script is not run anymore, since we now use automatic outliers detection. For documentation only.

```{r eval=FALSE, echo=FALSE}
# plates 1-6
JE2.outliers <- tibble(sample_id = "JE2",
                       well_id = c("exp191121_4_B06", "exp191121_4_D06"))
df_outliers_plates1_6 <- read_csv("Data_analysis/outliers/PI_kinetics_VANANZ_outliers_plates1_6.csv") %>%
  bind_rows(JE2.outliers)

# plates M1-M2
outliers_platesM1_M2 <- tibble(well_id = c(str_c("exp200309_M1_", c("G11", "C04", "B04", "G10", "E12")),
                               c("exp200313_M2_E04", "exp200313_M2_D12"))) %>%
  left_join(all_c_data %>% select(well_id, sample_id) %>% distinct())

# all outliers
df_outliers <- df_outliers_plates1_6 %>%
  bind_rows(outliers_platesM1_M2)
df_outliers

# save
write_csv(df_outliers,
          "Data_analysis/outliers/PI_data_VANANZ_outliers.csv")
```

## Concatenated dataframes, outliers removed manually

```{r eval=FALSE, echo=FALSE}
all_c_data_outliers_removed <- all_c_data %>%
  filter(!well_id %in% df_outliers$well_id)

# save
# write_csv(all_c_data_outliers_removed,
#           "Data_analysis/dataframes/concatenated_checked/all_cell_death_data_outliers_manually_removed.csv")
```

