---
title: "Generate datasets of within host pairs"
author: "Stefano Giulieri"
date: "30/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(paste0(here::here(), "/6-InToxSa_paper/Within_host_pairs"))
getwd()
```

Here we get PI uptake and cytoxicity data for within host pairs that were investigated in Giulieri, Genome Med 2018

```{r}
library(tidyverse)
library(patchwork)
rm(list = ls())
```

# List of within-host pairs

```{r}
within_host <- read_csv("../../plate_info/strain_metadata.csv") %>%
  filter(included)
```

Add patient IDs used in Giulieri, Genome Medicine 2018

```{r}
patient_id <- read_csv("~/OneDrive - The University of Melbourne/R/VANESSA_ANZCOSS/dataframes/whost_n130_df_new.csv") %>% select(strain_group = studyid, strain_group2 = patient_code_n130) %>%
  distinct()

within_host <- within_host %>%
  inner_join(patient_id)
```


How many have supernatant cytoxicity data?

```{r}
f <- "~/OneDrive - The University of Melbourne/R/VANESSA_ANZCOSS/data/whost_toxicity.csv"
file.copy(f, "raw_data")
cytox_supernat <- read_csv("raw_data/whost_toxicity.csv") %>%
  rename(sample_id = mdu_id, pct_dead_cells_range = range)
```

Get PI uptake parameters

```{r}
PI_data <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/PI/dataframes/parameters/PI_sample_parameters.Rda") %>%
  inner_join(within_host)
```

One merged dataframe

```{r}
PI_within_host <- within_host %>%
  inner_join(PI_data) %>%
  inner_join(cytox_supernat) %>%
  group_by(strain_group) %>%
  filter(n_distinct(sample_id) > 1) %>%
  ungroup()
```

Prepare for plotting

```{r}
df_plot <- PI_within_host %>%
  group_by(strain_group) %>%
  mutate(strain_group = fct_reorder(strain_group, AUC_death_mean, .desc = T)) %>%
  mutate(strain_status = if_else(intrahost_index == 1, "Baseline", "Evolved")) 
```

Change when evolved = colonising

```{r}
df_plot %>%
  filter(strain_status == "Evolved" & sample_type == "paired-colonising") %>%
  select(sample_id, strain_group, strain_status, sample_type, intrahost_sampledelay)

new_status <- df_plot %>%
  filter(strain_status == "Evolved" & sample_type == "paired-colonising") %>%
  group_by(strain_group) %>%
  slice_min(intrahost_sampledelay) %>%
  select(sample_id, strain_group) %>%
  mutate(new_status = "Baseline")

df_plot <- df_plot %>%
  mutate(patient_strain_group = strain_group) %>%
  mutate(strain_status = case_when(
    strain_group %in% new_status$strain_group & sample_id %in% new_status$sample_id ~ "Baseline",
    strain_group %in% new_status$strain_group & !sample_id %in% new_status$sample_id ~ "Evolved",
    TRUE ~ strain_status
  )) %>%
  mutate(sample_id = fct_reorder(sample_id, strain_status))
```


Plot these data

Panel A: PI uptake

```{r}
p1 <- df_plot %>%
  ggplot(aes(x = sample_id, y = AUC_death_mean, fill = strain_status)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = AUC_death_mean - AUC_death_sd, ymax = AUC_death_mean + AUC_death_sd)) +
  # geom_rect(data = df_plot %>% filter(strain_group %in% c("P02", "P05", "P12", "P17", "P20")), 
  #           xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA, colour = "red", size = 2) +
  facet_grid(~strain_group2, scales = "free_x", space = "free_x") +
  scale_fill_manual(values = c("#fee0d2", "#de2d26"), name = "", labels = c("Baseline", "Evolved")) +
  labs(x = "", y = "PI uptake AUC") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
 p1 
```

Panel B: cyotox supernatant

```{r}
p2 <- df_plot %>%
  ggplot(aes(x = sample_id, y = pct_dead_cells, fill = strain_status)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = pct_dead_cells - pct_dead_cells_range, ymax = pct_dead_cells + pct_dead_cells_range)) +
  # geom_rect(data = df_plot %>% filter(strain_group == "P20"), 
  #           xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA, colour = "blue", size = 2) +
  facet_grid(~strain_group2, scales = "free_x", space = "free_x") +
  scale_fill_manual(values = c("#deebf7", "#3182bd"), name = "", labels = c("Baseline", "Evolved")) +
  labs(x = "", y = "% dead cells") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
  
p2
```

Combine 

```{r}
p1 / p2 +
  plot_annotation(tag_levels = "A")

# ggsave("figures/within_host_pairs_toxicity_barplots.pdf")
```

# Stats

```{r}
df_stat <- df_plot %>%
  filter(strain_group == "VSS-541") %>%
  select(sample_id, pct_dead_cells, pct_dead_cells_range) %>%
  mutate(xmin = pct_dead_cells - pct_dead_cells_range/2,
         xmax = pct_dead_cells + pct_dead_cells_range/2) %>%
  select(sample_id, xmin, xmax) %>%
  pivot_longer(cols = c(xmin, xmax), names_to = "rank", values_to = "x")

wilcox.test(x ~ sample_id, data = df_stat)
t.test(x ~ sample_id, data = df_stat)
oneway.test(x ~ sample_id, data = df_stat)

```

This shows that a non-parametric test can't be used here because of the low number of observations.

T test works, however, normal distribution can't be assessed here! We prefer the one-way analysis of the variance (ANOVA) because some pairs have more than two strains. For ANOVA see https://statsandr.com/blog/anova-in-r/

```{r}
df_stats <- df_plot %>%
  mutate(xmin = pct_dead_cells - pct_dead_cells_range/2,
         xmax = pct_dead_cells + pct_dead_cells_range/2) %>%
  select(strain_group, strain_group2, sample_id, xmin, xmax) %>%
  pivot_longer(cols = c(xmin, xmax), names_to = "rank", values_to = "x") 

test_list <- list()
p_value <- c()

for (i in unique(df_stats$strain_group2)){
  df <- df_stats %>%
    filter(strain_group2 == i)
  out <- oneway.test(x ~ sample_id, data = df)
  print(out)
  test_list[[i]] <- out
  p_value[i] <- out$p.value
}

# Bonferroni correction
p_value <- p.adjust(p_value, method = "bonferroni")
```

List of pairs with significant difference
<<<<<<< HEAD

```{r}
supernat_signif <- p_value[p_value < .05]
```

Note that there is a 'tidier' approach to this using `dplyr` (faster, cleaner, more elegant). If we have time, we will rewrite this code using `dplyr` verbs and the `broom` package.

Now we repeat the test for PI uptake values

First, we need all replicate measurement rather than summary values

```{r}
PI_replicates <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/PI/dataframes/parameters/PI_parameters.Rda") %>%
  inner_join(within_host)

df_stats <- PI_replicates %>%
  select(sample_id, strain_group, strain_group2, AUC_death) 

test_list <- list()
p_value <- c()

for (i in unique(df_stats$strain_group2)){
  df <- df_stats %>%
    filter(strain_group2 == i)
  out <- oneway.test(AUC_death ~ sample_id, data = df)
  print(out)
  test_list[[i]] <- out
  p_value[i] <- out$p.value
}

# Bonferroni correction
# Bonferroni correction
p_value <- p.adjust(p_value, method = "bonferroni")

tibble(strain_group = names(p_value),
       p_value = p_value) %>%
  ggplot(aes(x = fct_reorder(strain_group, p_value), y = -log10(p_value))) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = -log10(.05)) +
  coord_flip()
```

=======

```{r}
supernat_signif <- p_value[p_value < .05]
```

Note that there is a 'tidier' approach to this using `dplyr` (faster, cleaner, more elegant). If we have time, we will rewrite this code using `dplyr` verbs and the `broom` package.

Now we repeat the test for PI uptake values

First, we need all replicate measurement rather than summary values

```{r}
PI_replicates <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/PI/dataframes/parameters/PI_parameters.Rda") %>%
  inner_join(within_host)

df_stats <- PI_replicates %>%
  select(sample_id, strain_group, strain_group2, AUC_death) 

test_list <- list()
p_value <- c()

for (i in unique(df_stats$strain_group2)){
  df <- df_stats %>%
    filter(strain_group2 == i)
  out <- oneway.test(AUC_death ~ sample_id, data = df)
  print(out)
  test_list[[i]] <- out
  p_value[i] <- out$p.value
}

# Bonferroni correction
# Bonferroni correction
p_value <- p.adjust(p_value, method = "bonferroni")

tibble(strain_group = names(p_value),
       p_value = p_value) %>%
  ggplot(aes(x = fct_reorder(strain_group, p_value), y = -log10(p_value))) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = -log10(.05)) +
  coord_flip()
```

>>>>>>> b7d90018bba4a5f6b8bc9bc55a272d96fa14f1ee
List of pairs with significant difference

```{r}
pi_signif <- p_value[p_value < .05]
```

New figure

```{r}
p1 <- p1 +
  geom_rect(data = df_plot %>% filter(strain_group2 %in% names(pi_signif)),
            xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA, colour = "red", size = 2)
<<<<<<< HEAD

p2 <- p2 +
  geom_rect(data = df_plot %>% filter(strain_group2 %in% names(supernat_signif)),
            xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA, colour = "blue", size = 2)

p2 / p1 +
  plot_annotation(tag_levels = "A")

ggsave("figures/within_host_pairs_toxicity_barplots_stats.pdf")
=======
>>>>>>> b7d90018bba4a5f6b8bc9bc55a272d96fa14f1ee
```

