---
title: "Concatenate and check growth data"
author: "Stefano Giulieri"
date: "12/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This is a script to concatenate all growth curves dataframes in a single file and then perform data checks. 


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
msg <- stringr::str_c("My directory is ", here::here())
message(msg)
```

```{r}
library(tidyverse)
library(magrittr)
rm(list = ls())
source("Functions/all_functions.R")
```

# Concatenate single parsed dataframes 

As of 12 October 2020, we have now parsed growth curve data as single experiments that can now be concatenated.

## List of available parsed growth curves dataframes

```{r}
g <- list.files("Data_parsing/parsed_experiments", recursive = T, full.names = TRUE) %>% str_subset("growth")
g
```

We have a total of `r length(g)` experiments with growth data. However, not all experiments will retained. We exclude experiments with less than 40k cells

## Concatenate raw data

```{r}
# all cell death data
all_g_data_raw <- purrr::map(g, function(g){
  df <- read_csv(g) %>%
    mutate(plate = as.character(plate))
}) %>%
  bind_rows()
# experiments with less than 40k cells
cell_low <- c("exp190905", 
              "exp190909",
              "exp190921",
              "exp190923")
all_g_data <- all_g_data_raw %>%
  filter(!experiment %in% cell_low)
```

## Check concatenated file

```{r}
# included plates / experiments
all_g_data %>%
  select(plate, experiment) %>%
  distinct() %>%
  knitr::kable(row.names = T)
```

Two experiments (6 replicates) for each of plates 2, 3 and 4 and GP1, GP2, GP4-GP7 and one experiment (one replicate) for each of plates 1, 5, 6, P1, P2, M1, M2, GP3 (one of the GP3 experiments failed). In total 26 experiments.

```{r}
# Number of wells per experiment
all_g_data %>%
  group_by(plate, experiment) %>%
  summarise(n_wells = n_distinct(well)) %>%
  knitr::kable(row.names = T)
```

```{r}
# Missing data
all_g_data %>%
  filter(is.na(OD)) %>%
  count(plate, experiment, sample_id)
```


We have now confirmed that the concatenated dataframe includes all experiments of interest and the all wells have been parsed.

```{r}
write_csv(all_g_data,
          "Data_analysis/dataframes/concatenated_checked/growth_data_parsed.csv")
```
