---
title: "Description and analysis of genetic pairs phylogenies"
author: "Stefano Giulieri"
date: '28/03/2023'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/8-eLife_revision/Genetic_pairs"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we extract phylogenetic data from each genetic pair: MRCA, chain of ancestors (nodes) connecting iso1 to iso2, all tips of the smallest clade containing iso1 and iso2 (descendants/offspring of the MRCA)

```{r}
library(tidyverse)
library(ggtree)
library(tidytree)
library(tictoc)
library(furrr)

rm(list = ls())
```


# Raw data

Pairs

```{r}
pairs <- readRDS("processed_data/PI_pairs_data/pairs_description_dataframe.Rda")
```

Tree

```{r}
tree1 <- read.tree("raw_data/tree/iqtree.treefile")

# tree2 <- read.tree("../../Genetic_pairs_analysis_Oct_2020/metadata/iqtree.treefile")
```

We keep only tips with PI information

```{r}
tree <- ape::keep.tip(tree1, pairs$iso1) %>%
  phytools::midpoint.root() 
ggtree(tree)
ape::write.tree(tree, "processed_data/tree/iqtree_PI_tips.tree")
```

# Extract clade information 

For each pair extract MRCA, number of descendants, MRCA level

```{r}
tree_data <- as.treedata(tree)
df_tree <- as_tibble(tree_data)
pair <- pairs[1,]
x <- pair$iso1
x <- which(tree$tip.label == x)
y <- pair$iso2
y <- which(tree$tip.label == y)

mrca <- MRCA(tree, c(x, y))

x1 <- ancestor(tree, x)
y1 <- ancestor(tree, y)
d <- c(setdiff(x1, y1), mrca, rev(setdiff(y1, x1)))
d <- c(setdiff(x1, y1), rev(setdiff(y1, x1)))
d1 <- which(x1 == mrca)
d2 <- which(y1 == mrca)
D <- d1 + d2

DD <- c(d, x, y)
p <- ggtree(tree) +
  geom_tree(aes(colour = node %in% DD)) +
  geom_tiplab(aes(subset = label %in% c(pair$iso1, pair$iso2))) +
  geom_point2(aes(subset = node == mrca)) +
  scale_colour_manual(values = c("grey", "blue"), guide = "none")
p
# pp <- ggplot_build(p)
# 
# 
# pp1 <- layer_data(p)
# pp2 <- layer_data(p, 2)
# pp3 <- layer_data(p, 3)
# pp4 <- layer_data(p, 4)





# plot(tree, use.edge.length = F, show.tip.label = F, edge.color = if_else(tree$edge[,2] %in% c(d, x, y), "blue", "gray" ), no.margin = T)


o <- tidytree::offspring(tree, mrca, tiponly = T)
O <- length(o)

```

Write a function to do all of this

```{r}
get_ancestral_path <- function(tree, iso1, iso2, return_plot = F){
  x <- which(tree$tip.label == iso1)
  y <- which(tree$tip.label == iso2)

  mrca <- MRCA(tree, c(x, y))

  x1 <- ancestor(tree, x)
  y1 <- ancestor(tree, y)
  d1 <- which(x1 == mrca)
  d2 <- which(y1 == mrca)
  d <- c(setdiff(x1, y1), rev(setdiff(y1, x1)))
  D <- d1 + d2

  DD <- c(d, x, y)
  
  tips <- c(iso1, iso2)
  
  p <- ggtree(tree) +
  geom_tree(aes(colour = node %in% DD)) +
  geom_tiplab(aes(subset = label %in% {{tips}})) +
  geom_point2(aes(subset = node == mrca)) +
  scale_colour_manual(values = c("grey", "blue"), guide = "none")
  
  o <- tidytree::offspring(tree, mrca, type = "tips")
  O <- length(o)
  
  out <- list(iso1 = iso1,
              iso2 = iso2,
              mrca = mrca, 
              path_nodes = d,
              path_length = D,
              offspring_tips = o,
              offspring_labels = df_tree %>% filter(node %in% o) %>%.$label,
              n_offspring = O)
  
  if (return_plot) out$plot <- p
  return(out)
}

out <- get_ancestral_path(tree, pair$iso1, pair$iso2)
out <- get_ancestral_path(tree, pair$iso1, pair$iso2, return_plot = T)
out$plot

```

Get mrca for each pair

```{r}
tic()
pairs_with_tree_data <- pairs[1:10,] %>%
  group_by(pair_id) %>%
  nest() %>%
 mutate(ancestral_path = purrr::map(data, ~get_ancestral_path(tree, .x$iso1, .x$iso2))) %>%
  mutate(mrca = map_int(ancestral_path, "mrca"),
         path_len = map_int(ancestral_path, "path_length"),
         offspring_tips = map(ancestral_path, "offspring_tips"),
         n_offspring = map_int(ancestral_path, "n_offspring")) %>%
  unnest(cols = data)
toc()

plan(multisession, workers = 4)


tic()
get_ancestral_path_out <- pairs %>%
  group_split(pair_id) %>%
  future_map(~get_ancestral_path(tree, .x$iso1, .x$iso2), .progress = T) 
toc()
# 9298.983 sec elapsed

df_get_ancestral_path <- get_ancestral_path_out %>%
  do.call(rbind, .) %>%
  as_tibble() 


df_get_ancestral_path <- df_get_ancestral_path %>%
  unnest(c(iso1, iso2, mrca, path_length, n_offspring))
```

Save intermediate stages. Because these large files are problematic in github, we save them in the local copy of the directory 

```{r}
# get_ancestral_path_out %>%
#   saveRDS("processed_data/PI_pairs_data/get_ancestral_path_list.Rda")
# df_get_ancestral_path %>%
#   saveRDS("processed_data/PI_pairs_data/get_ancestral_path_dataframe.Rda")
```


