---
title: "Generate dataframe of pairs information and summary plots of pairs metadata"
author: "Stefano Giulieri"
date: '20/02/2023'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/8-eLife_revision/Genetic_pairs"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we extract pairs with increase of PI uptake from iso1 to iso2 that are 1) *not already included in the 28 genetic pairs* ; 2) phylogenetically independent.

```{r}
library(tidyverse)

rm(list = ls())
```

# Raw data

Pairs data collected so far

```{r}
pairs_data <- readRDS("processed_data/PI_pairs_data/pairs_description_with_phylo_dataframe.Rda")

mutations <- readRDS("processed_data/variants_combined_annotation/snippy_denovo_annotated.Rda")
```

Variables to add: PAIR_ID, ST, CC

# Visualisations

```{r}
pairs_data %>%
  ggplot(aes(x = h_ratio, y = delta_auc)) +
  geom_point() +
  facet_wrap(~PI_pair)

# Manual downsampling
df1 <- pairs_data %>%
  ungroup() %>%
  filter(!PI_pair) %>%
  slice_sample(prop = .01)
df2 <- pairs_data %>%
  ungroup() %>%
  filter(PI_pair) 
df_plot <- bind_rows(df1, df2)


# Now represent genetic distance and pheno distance on main axes

p1 <- df_plot %>%
  ggplot(aes(x = dist_mash, y = delta_auc, shape = PI_pair, colour = log_p)) +
  geom_point() +
  stat_ellipse(aes(alpha = PI_pair), geom = "path", show.legend = T)+
    scale_alpha_discrete( name = "Gen pair", labels = c("No", "Yes")) +
  scale_shape_discrete( name = "Gen pair", labels = c("No", "Yes")) +
   scale_color_gradient(low = "blue", high = "yellow", name = "-log(p)") +
  labs(x = "Genetic distance", y  = "Difference in PI uptake") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"))
p1
```

The pairs of interest are in the top left corner of the plot

```{r}
p2 <- pairs_data %>%
  filter(dist_denovo <= 200 & pval < .05) %>%
  ggplot(aes(x = dist_mash, y = delta_auc, shape = PI_pair, colour = log_p)) +
  geom_point() +
   scale_color_gradient(low = "blue", high = "yellow", name = "-log(p)") +
  labs(x = "Genetic distance", y  = "Difference in PI uptake") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"))
p2
```

Now we try to define 

```{r}
df_gp_lp <- pairs_data %>%
  filter(dist_denovo <= 200 & pval < .05)
```

Exclude all pairs that overlap with with 28 genetic pairs (the clade shares descendants / offspring)

```{r}
all_offspring <- pairs_data %>%
  filter(PI_pair) %>%
  .$offspring_tips %>%
  unlist() %>%
  unique()
df_gp_lp <- df_gp_lp %>%
  rowwise() %>%
  filter(!any(offspring_tips %in% all_offspring))
```

Keep only positive pairs

```{r}
df_gp_lp_highPI <- df_gp_lp %>%
  filter(delta_auc > 30) %>%
  mutate(pheno_geno_score = delta_auc/dist_denovo) %>%
  arrange(-pheno_geno_score) %>%
  ungroup()
```

Now select pairs that are independent from each other

```{r}
source("../../0-Functions/all_functions.R")
df_highPI <- generate_gp_replicates(df = df_gp_lp_highPI)
```

Which mutations

```{r}
df_highPI_mutations <- df_highPI %>%
  inner_join(mutations) %>%
  relocate(neb_gene_symbol, MUTATION_SHORT)
```

Quick look at convergence

```{r}
df_converg <- df_highPI_mutations %>%
  drop_na(cdhit_group) %>%
  filter(EFFTYPE_SHORT %in% c("truncating", "non-synonymous")) %>%
  group_by(cdhit_group, neb_gene_symbol, neb_product, neb_locus_tag) %>%
  summarise(n_events = n_distinct(pair_id)) %>%
  arrange(-n_events)
```

Now including the old pairs

```{r}
df_pairs <- pairs_data %>%
  filter(PI_pair) %>%
  bind_rows(df_highPI)

df_diffPI_mutations <- df_pairs %>%
  inner_join(mutations) %>%
  relocate(neb_gene_symbol, MUTATION_SHORT)

df_converg2 <- df_diffPI_mutations %>%
  drop_na(cdhit_group) %>%
  filter(EFFTYPE_SHORT %in% c("truncating", "non-synonymous")) %>%
  group_by(cdhit_group, neb_gene_symbol, neb_product, neb_locus_tag, neb_gene_length) %>%
  mutate(n_events = n_distinct(pair_id)) %>%
  arrange(-n_events, neb_gene_symbol) %>%
  relocate(n_events)
```

# Save processed data

```{r}
df_converg2 %>%
  saveRDS("processed_data/gene_convergence/PI_discordant_with_PI_increase/df_PI_convergence_analysis_diffPI.Rda")
```

Dataframe with genes with >= 2 independent mutations. Exclude genes with no FPR3757 homolog and exclude ebh

```{r}
df_converg_genes <- df_converg2 %>%
  filter(n_events > 1) %>%
  filter(neb_gene_symbol != "ebh") %>%
  drop_na(neb_locus_tag) %>%
  distinct()

# stats
df_converg_genes %>%
  distinct(neb_gene_symbol, n_events) %>%
  ungroup() %>%
  count(n_events)

df_converg_genes %>%
   saveRDS("processed_data/gene_convergence/PI_discordant_with_PI_increase/df_PI_convergent_genes_diffPI.Rda")
```
