---
title: "Generate dataframe of pairs information and summary plots of pairs metadata"
author: "Stefano Giulieri"
date: '17/06/2022'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c("~/Documents/Github/InToxSa", "/8-eLife_revision/Genetic_pairs"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we put together all pairs metadata (phenotype, genetic distance, ST, CC, phylo data). Using these combined dataframe, we try to find plots to summarise metadata for both ~140k pairs with available metadata and 28 pairs with contrasting PI.

```{r}
library(tidyverse)

rm(list = ls())
```

# Raw data

Pairs data collected so far

```{r}
pairs_data <- readRDS("processed_data/PI_pairs_data/pairs_description_with_phylo_dataframe.Rda")
```

Variables to add: PAIR_ID, ST, CC

# Visualisations

```{r}
pairs_data %>%
  ggplot(aes(x = h_ratio, y = delta_auc)) +
  geom_point() +
  facet_wrap(~PI_pair)

# Manual downsampling
df1 <- pairs_data %>%
  ungroup() %>%
  filter(!PI_pair) %>%
  slice_sample(prop = .01)
df2 <- pairs_data %>%
  ungroup() %>%
  filter(PI_pair) 
df_plot <- bind_rows(df1, df2)

df_plot %>%
  ggplot(aes(x = delta_auc, y = log_p,  colour = PI_pair)) +
  geom_point()+
  stat_ellipse() +
  facet_wrap(~h_ratio > .5) +
  scale_colour_manual(values = c("red", "blue")) +
    theme_bw()

df_plot %>%
  ggplot(aes(x = delta_auc, y = log_p,  shape = PI_pair )) +
  geom_point(aes(colour = h_ratio))+
  stat_ellipse(aes(fill = PI_pair)) +
  scale_fill_manual(values = c("red", "blue")) +
  scale_colour_viridis_c() +
  facet_wrap(~PI_pair) +
    theme_bw()

df_plot %>%
  ggplot(aes(x = delta_auc, y = h_ratio,  colour = PI_pair)) +
  geom_point()+
  stat_ellipse() +
  scale_colour_manual(values = c("red", "blue")) +
    theme_bw()
```

Now represent genetic distance and pheno distance on main axes

```{r}
p1 <- df_plot %>%
  ggplot(aes(x = dist_mash, y = delta_auc, shape = PI_pair, colour = log_p)) +
  geom_point() +
  stat_ellipse(aes(alpha = PI_pair), geom = "path", show.legend = T)+
    scale_alpha_discrete( name = "Gen pair", labels = c("No", "Yes")) +
  scale_shape_discrete( name = "Gen pair", labels = c("No", "Yes")) +
   scale_color_gradient(low = "blue", high = "yellow", name = "-log(p)") +
  labs(x = "Genetic distance", y  = "Difference in PI uptake") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"))
p1
```

Alternative figure after discussing with Tim, Abdou and Calum

```{r}
p1a <- df_plot %>%
  ggplot(aes(x = dist_mash, y = delta_auc,  colour = PI_pair)) +
  geom_point() +
  stat_ellipse(aes(colour = PI_pair, fill = PI_pair), geom = "polygon", alpha = .3, show.legend = F)+
    scale_colour_manual( values = c("red", "blue"), name = "Gen pair", labels = c("No", "Yes")) +
  labs(x = "Genetic distance", y  = "Difference in PI uptake") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"))
p1a
```


Focus on PI pairs only

```{r}
df_plot <- df_plot %>%
  filter(PI_pair)

df_plot %>%
  ggplot(aes(x = h_ratio, y = delta_auc)) +
  geom_point()

df_plot %>%
  ggplot(aes(x = dist_denovo, y = delta_auc, colour = h_ratio)) +
  geom_point()

df_plot %>%
  ggplot(aes(x = dist_denovo, y = path_length)) +
  geom_point()

df_plot %>%
  ggplot(aes(x = dist_denovo, y = delta_auc, size = path_length)) +
  geom_point()

df_plot %>%
  ggplot(aes(x = dist_denovo, y = delta_auc, colour = path_length)) +
  geom_point()
```

Ballon plot

```{r}
df_plot %>%
  ggplot(aes(x = fct_reorder(pair_id, delta_auc, .desc = T), y = delta_auc, size = dist_denovo, colour = path_length)) +
  geom_point() +
  coord_flip() +
  theme_bw()

p2 <- df_plot %>%
  ggplot(aes(x = fct_reorder(pair_id, dist_denovo, .desc = F), y = dist_denovo, size = path_length, colour = delta_auc)) +
  geom_point() +
  scale_colour_continuous(high = "#efedf5", 
                          low = "#3f007d", name = "Delta PI uptake") +
  scale_size_continuous(name = "Phylo distance") +
  coord_flip() +
  labs(y = "# mutations", x = "Pair ") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"),
        legend.position = c(.8,.5))
p2
```

Changes suggested by Tim, Abdou and Calum

```{r}
p2a <- df_plot %>%
  ggplot(aes(x = fct_reorder(pair_id, dist_denovo, .desc = F), y = dist_denovo, colour = delta_auc)) +
  geom_point(size = 3) +
  scale_colour_continuous(high = "#efedf5", 
                          low = "#3f007d", name = "Delta PI uptake") +
  scale_size_continuous(name = "Phylo distance") +
  coord_flip() +
  labs(y = "# mutations", x = "Pair ") +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"),
        legend.position = c(.8,.5))
p2a
```


# Save processed data

```{r}
# list(p1, p1a, p2, p2a) %>%
#   saveRDS("processed_data/PI_pairs_data/pairs_metadata_plots.Rda")
```

