---
title: 'Generate figures for the convergent evolution analysis'
author: "Stefano Giulieri"
date: "17/06/2022"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/InToxSa_paper/Genetic_pairs"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```


Here we generate figure for the gene enrichment analysis. The whole gene convergence analysis process is described in `Gene_convergence_analysis.Rmd`.

To be consistent with the previous work we use the same dataset generated from the previous analysis

```{r}
library(tidyverse)
library(patchwork)
library(ggrepel)
library(cols4all)
rm(list = ls())
```

# Load data

```{r}
df_mutations <- readRDS("raw_data/df_PI_convergent_genes.Rda")
```

NEB kinetics data

```{r}
df_PI_NEB <- readRDS("raw_data/PI_sample_parameters_plateGPV1.Rda")

df_PI_NEB_wells <- readRDS("raw_data/PI_parameters_plateGPV1.Rda")
```

# Figure 3: gene enrichment in the entire dataset

```{r}
base_sz <- 20
point_sz <- 3
colors <- RColorBrewer::brewer.pal(n = 6, name = "PuOr")
names(colors) <- rev(c("IS insertion (intragenic)", "IS insertion (intergenic)", "large deletion", "intergenic", "non-synonymous", "truncating"))
colors_genes <- colors[c(1,2,6)]
colors_genes
```

## Panel A 

### Mutation counts

```{r}
df_plot <- df_mutations %>%
   group_by(cdhit_group, cdhit_symbol,EFFTYPE_SHORT) %>%
  summarise(n_mutations = n_distinct(pair_id)) 


# set an order to be consistent across plots
cluster_order <- df_plot %>%
  group_by(cdhit_group) %>%
  summarise(n_mutations = sum(n_mutations)) %>%
  arrange(desc(n_mutations), cdhit_group) %>%
  transmute(cdhit_group,
            cluster_order = row_number())

# Optional filtering
cluster_order <- cluster_order %>%
  filter(cluster_order <= 20)

df_plot <- df_plot %>%
  right_join(cluster_order) 


# Plot A
p1 <- df_plot %>%
  ggplot(aes(x = fct_reorder(cdhit_symbol, cluster_order, .desc = T),
             y = n_mutations,
             fill = EFFTYPE_SHORT)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks = c(0,5,10), minor_breaks = c(1:4,6:9)) +
 scale_fill_manual(values = colors, name = "") +
  labs(x = "", y = "# mutations") +
  coord_flip() +
  theme_bw() +
  theme(text = element_text(face = "bold"))

p1
```

Same but with CC

```{r}
df_plot <- df_mutations %>%
   group_by(cdhit_group, cdhit_symbol, CC_iso1,EFFTYPE_SHORT) %>%
  summarise(n_mutations = n_distinct(pair_id)) 


# set an order to be consistent across plots
cluster_order <- df_plot %>%
  group_by(cdhit_group) %>%
  summarise(n_mutations = sum(n_mutations)) %>%
  arrange(desc(n_mutations), cdhit_group) %>%
  transmute(cdhit_group,
            cluster_order = row_number())

# Optional filtering
cluster_order <- cluster_order %>%
  filter(cluster_order <= 20)

df_plot <- df_plot %>%
  right_join(cluster_order) 


# Plot A
p1 <- df_plot %>%
  ggplot(aes(x = fct_reorder(cdhit_symbol, cluster_order, .desc = T),
             y = n_mutations,
             fill = as.factor(CC_iso1))) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks = c(0,5,10), minor_breaks = c(1:4,6:9)) +
cols4all::scale_fill_discrete_c4a_cat(palette = "tol.muted", name = "CC") +
  labs(x = "", y = "# mutations") +
  coord_flip() +
  theme_bw() +
  theme(text = element_text(face = "bold"))

p1
```


# Plot B: mutations mini-maps

Here we map the position of the mutation within the coding region. Substitutions, truncations and IS insertions are designed by the position (or start position), large deletions by the start position (if the whole gene is deleted the deletion is shown at the first codon).

Plot B has to follow the same gene order as plot A. 

```{r}
df_gene_maps <- df_mutations %>%
  right_join(cluster_order) %>%
  group_by(cdhit_group, cdhit_symbol, AA_POS, EFFTYPE_SHORT, n_events, cluster_order) %>%
  summarise(n_events = n_distinct(pair_id)) 

df_rect <- df_mutations %>%
  right_join(cluster_order) %>%
  group_by(cdhit_group, n_events, cluster_order) %>%
  summarise(xmin = cluster_order - 0.25,
            xmax = cluster_order + 0.25, 
            ymin = 0,
            ymax = str_length(cdhit_seq))
  

p2 <- df_gene_maps %>%
  ggplot(aes(x = cluster_order, y = AA_POS, colour = EFFTYPE_SHORT)) +
  geom_point() +
  geom_rect(data = df_rect, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            inherit.aes = F, alpha = 0, color = "black") +
  scale_y_continuous(limits = c(0,2500), breaks = c(0,1000, 2000)) +
  scale_x_reverse(expand = c(0, .4)) +
  scale_color_manual(values = rev(colors), na.translate = F, guide = F) +
  # scale_size(breaks = 1:3, name = "N independent mutations", guide = F) +
  coord_flip() +
  labs(x = "") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom") +
  theme(text = element_text(face = "bold"))

p2
```

# Panel C: Nebraska

For each mutant, calculate the mean AUC change (+/- SD) as compared to JE2

```{r}
df_neb_stats <- df_PI_NEB %>%
  select(sample_id, contains("AUC")) %>%
  left_join(df_mutations %>% distinct(cdhit_group, neb_mutant_id),
            by = c("sample_id" = "neb_mutant_id"))

wt_pvalue <- c()
for (i in unique(df_PI_NEB_wells$sample_id)){
 x <- df_PI_NEB_wells %>%
    filter(sample_id == i) %>%
   .$AUC_death
 y <- df_PI_NEB_wells %>%
    filter(sample_id == "JE2") %>%
   .$AUC_death
 wt <- wilcox.test(x, y)
 wt_pvalue[i] <- wt$p.value
}

df_neb_stats <- df_neb_stats %>%
  left_join(tibble(sample_id = names(wt_pvalue),
                   pvalue = wt_pvalue)) %>%
  mutate(pvalue_corr = p.adjust(pvalue, method = "fdr"))

# p values distribution
df_neb_stats %>%
  ggplot(aes(x = -log10(pvalue))) +
  geom_histogram() +
  geom_density(aes(y =  ..count..))

# check identical values: this is normal with wilcox test, especially at very low values
# "All the usual rank based tests have a discrete distribution based on counting of arrangements."
# https://stats.stackexchange.com/questions/232927/same-p-value-is-it-normal


AUC_JE2 <- df_neb_stats %>% filter(sample_id == "JE2") %>% .$AUC_death_mean
df_neb_stats %>%
  ggplot(aes(x = fct_reorder(sample_id, AUC_death_mean),
             y = AUC_death_mean,
             ymin = AUC_death_mean - AUC_death_sd,
             ymax = AUC_death_mean + AUC_death_sd,
             colour = -log10(pvalue_corr))) +
  geom_point() +
  geom_linerange() +
  geom_hline(yintercept = AUC_JE2, linetype = "dashed") +
  scale_colour_viridis_c() +
  coord_flip() +
  theme_bw()

df_neb_stats %>%
  ggplot(aes(x = fct_reorder(sample_id, AUC_death_mean),
             y = AUC_death_mean,
             ymin = AUC_death_mean - AUC_death_sd,
             ymax = AUC_death_mean + AUC_death_sd,
             colour = pvalue_corr < .05)) +
  geom_point() +
  geom_linerange() +
  geom_hline(yintercept = AUC_JE2, linetype = "dashed") +
  scale_colour_manual(values = c("red", "blue")) +
  coord_flip() +
  theme_bw()

df_plot <- df_neb_stats %>%
  right_join(cluster_order)

p3 <-  df_plot %>%
  ggplot(aes(x = -cluster_order,
             y = AUC_death_mean,
             ymin = AUC_death_mean - AUC_death_sd,
             ymax = AUC_death_mean + AUC_death_sd,
             colour = pvalue_corr < .05)) +
  geom_point() +
  geom_linerange() +
  geom_hline(yintercept = AUC_JE2, linetype = "dashed") +
  scale_colour_manual(values = c("red", "blue"), guide = F) +
  labs(x = "", y = "PI uptake (AUC)") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

p3 
```

Final NEB plot

```{r}
df_plot <- df_neb_stats %>%
  right_join(cluster_order)

p3 <-  df_plot %>%
  ggplot(aes(x = -cluster_order,
             y = AUC_death_mean,
             ymin = AUC_death_mean - AUC_death_sd,
             ymax = AUC_death_mean + AUC_death_sd,
             colour = pvalue_corr < .05)) +
  geom_point() +
  geom_linerange() +
  geom_hline(yintercept = AUC_JE2, linetype = "dashed") +
  scale_colour_manual(values = c("blue", "skyblue"), guide = F) +
  labs(x = "", y = "PI uptake (AUC)") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(face = "bold"))

p3 
```

# Panel D: Operetta NEB

```{r}
Cell_count.df <- readRDS("raw_data/Cell_count_corrected.df.Rda") %>%
  mutate(cell_id = str_c(sample_id, plate_number, plate_replicate, replicate, timepoint, well, Field, Object_No, sep = "_"))

Spot_Cell_count.df <- readRDS("raw_data/Spot_Cell_count.df.Rda")

```

```{r}
# merge corrected bacteria count with cell count table
Cell_count_corrected.df <- merge(Cell_count.df, Spot_Cell_count.df, by = "cell_id", all.x = T) %>%
  select(-cor_nb_bacteria) %>%
  rename(cor_nb_bacteria = n) %>%
  mutate(cor_nb_bacteria = ifelse(is.na(cor_nb_bacteria), 0, cor_nb_bacteria)) 

Cell_count_corrected.df <- Cell_count_corrected.df %>%
  mutate(timepoint = factor(.$timepoint, levels = c("3h", "24h")))

Cell_count_corrected.df <- Cell_count_corrected.df %>%
  mutate(plate_replicate_id = paste0(plate_number, "_", plate_replicate))
```


```{r}
df <- Cell_count_corrected.df %>%
    filter(plate_number == "GPV1")  %>%
    mutate(sample_id = ifelse(sample_id == "TOX-4", yes = "PAM-agrA", no = sample_id)) %>%
    filter(sample_id != "Non infected") %>%
  left_join(df_plot) %>%
 filter(!is.na(cluster_order) | sample_id == "JE2") %>%
  replace_na(list(cluster_order = 21))

df_plot4 <- df %>%
    mutate(Infected = ifelse(Number_of_Spots>0, yes = "yes", no = "no")) %>%
    group_by(sample_id, cluster_order, timepoint,Infected) %>% 
    summarise(count = n()) %>% 
    mutate(perc = count/sum(count)*100)

p4 <- ggplot(df_plot4 , aes(x = fct_reorder(sample_id, cluster_order, .desc = T), y = timepoint,  fill = perc )) +
    geom_tile(stat = "identity") +
    xlab("") +
    ylab("") +
    #scale_fill_manual(name = "Infected:", labels = c("no" , "yes"), values = c("#4575b4", "#d73027")) +
    coord_flip() +
    theme_bw() +
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(), 
          axis.text = element_text(size = 12), 
          plot.title = element_text(size = 16, face = 'bold'), 
          axis.title = element_text(face = 'bold')) + 
    scale_fill_gradient(low = "white", high = "steelblue", name = "% infected")
  p4
  
  df_plot5 <- df %>%
    filter(cor_nb_bacteria > 0) %>%
    group_by(sample_id, cluster_order, timepoint) %>% 
    summarise(bacteria_nb_per_cell = mean(cor_nb_bacteria)) 
  
  p5 <- ggplot(df_plot5 , aes(x = fct_reorder(sample_id, cluster_order, .desc = T), y = timepoint,  fill = bacteria_nb_per_cell )) +
    geom_tile(stat = "identity") +
    xlab("") +
    ylab("") +
   # ggtitle(exp) +
    theme(plot.title = element_text(size = 15)) +
    #scale_fill_manual(name = "Infected:", labels = c("no" , "yes"), values = c("#4575b4", "#d73027")) +
    coord_flip() +
    theme_bw() +
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(), 
          axis.text = element_text(size = 12), 
          plot.title = element_text(size = 16, face = 'bold'), 
          axis.title = element_text(face = 'bold'),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank() ) + 
    scale_fill_gradient(low = "white", high = "red", name = "SA/infected cell")
  p5
  
p6 <- p4 + p5 +
  plot_layout(guides = "collect")
p6
```


# Save processed data

```{r}
list(p1, p2, p3,p4, p5, p6) %>%
  saveRDS("processed_data/gene_convergence/convergence_plots.Rda")
```

