---
title: "Description of genetic pairs wih contrasting PI"
author: "Stefano Giulieri"
date: '2022-06-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/InToxSa_paper/Genetic_pairs"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we describe our dataset of genetic pairs with contrasting PI: mean AUC of PI uptake, genetic distance, phylogenetic tree.

```{r}
library(tidyverse)
library(broom)

rm(list = ls())
```


# Raw data

## Dataset of genetic pairs 

```{r}
pairs <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/snippy/df_all_pairs_distances_metadata.Rda") %>%
  as_tibble() 
```

Add PI uptake AUC mean

```{r}
PI_parameters <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/PI/dataframes/parameters/PI_parameters.Rda") %>%
  as_tibble() %>%
  select(sample_id, pi_auc = AUC_death)

PI_mean <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/PI/dataframes/parameters/PI_sample_parameters.Rda") %>%
  as_tibble() %>%
  select(sample_id, pi_auc = AUC_death_mean)

pairs_with_pi <- pairs %>%
  filter(iso1 %in% PI_parameters$sample_id & iso2 %in% PI_parameters$sample_id)
```

For each pair calculate delta AUC and p value

```{r}
df_pairs_with_pi_auc <- pairs_with_pi %>%
 nest_join(PI_parameters, by = c("iso1" = "sample_id"), name = "x") %>%
  nest_join(PI_parameters, by = c("iso2" = "sample_id"), name = "y") 

df_wt <- df_pairs_with_pi_auc %>%
  mutate(wt = map2(x, y, ~wilcox.test(.x$pi_auc, .y$pi_auc)),
         pval = map_dbl(wt, "p.value")) %>%
  select(-c(x, y, wt))

df_wt %>%
  ggplot(aes(x = -log10(pval))) +
  geom_density()

df_wt %>%
  ggplot(aes(x = -log10(pval))) +
  geom_histogram()

```

## Final dataset of genetic pairs with contrasting PI

```{r}
PI_pairs <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/convergence/PI/df_cleaned_pair_PI.Rda")
```

# Scatterplot of PI pairs

Here we generate a scatter plot showing the genetic distance on the x axis and the phenotypic distance on the y axis

```{r}
df_plot <- df_wt %>%
  inner_join(PI_mean, by = c("iso1" = "sample_id")) %>%
  rename(iso1_pi_auc = pi_auc) %>%
  inner_join(PI_mean, by = c("iso2" = "sample_id")) %>%
  rename(iso2_pi_auc = pi_auc) %>%
  mutate(delta_auc = iso2_pi_auc - iso1_pi_auc) %>%
  select(pair_id, iso1, iso2, delta_auc, pval, starts_with("dist")) %>%
  mutate(PI_pair = pair_id %in% PI_pairs$pair_id)

df_plot %>%
  filter(PI_pair) %>%
  ggplot(aes(x = dist, y = delta_auc)) +
  geom_point()

df_plot %>%
  ggplot(aes(x = delta_auc, y = -log10(pval), colour = dist)) +
  geom_point() +
  scale_colour_viridis_c() +
  facet_wrap(~PI_pair)

df_plot %>%
  ggplot(aes(x = delta_auc, colour = -log10(pval), y = -log10(dist))) +
  geom_point() +
  scale_colour_viridis_c() +
  facet_wrap(~PI_pair)

df_plot %>%
  filter(pval < .05) %>%
  ggplot(aes(x = delta_auc, colour = PI_pair, y = -log10(dist))) +
  geom_point() +
 scale_colour_manual(values = c("red", "blue")) +
  theme_bw()

df_plot %>%
  ggplot(aes(x = delta_auc, colour = PI_pair, y = -log10(dist))) +
  geom_point() +
  scale_colour_manual(values = c("red", "blue")) +
  theme_bw()

df_plot %>%
  ggplot(aes(x = delta_auc, colour = PI_pair, y = dist_mash)) +
  geom_point() +
  scale_colour_manual(values = c("red", "blue")) +
  theme_bw()
```

Try with box plot

```{r}
df_plot %>%
  ggplot(aes(x = PI_pair, y = delta_auc)) +
  geom_boxplot()
```

Try with density

```{r}
df_plot <- df_plot %>%
  mutate(log_p = -log10(pval),
         log_dist = log10(dist),
         log_dist_new = log10(dist_new),
         log_dist_cophenetic = log10(dist_cophenetic_new),
         log_dist_mash = -log10(dist_mash)) %>%
  relocate(c(pair_id, iso1, iso2, dist_denovo, PI_pair))

plot_list <- list()
for (v in colnames(df_plot)[-c(1:5)]){
  s <- rlang::sym(v)
  p <- df_plot %>%
  ggplot(aes(x = {{s}}, fill = PI_pair, colour = PI_pair)) +
 geom_density(alpha = .3) +
  theme_bw()
  plot_list[[v]] <- p
}
plot_list
```

# Save processed data

```{r}
plot_list %>%
  saveRDS("processed_data/pairs_description_density_plots.Rda")

df_plot %>%
  saveRDS("processed_data/pairs_description_dataframe.Rda")
```

