---
title: "Prepare figure 2"
author: "Stefano Giulieri"
date: "24/05/2022"
output:
  html_document:
    toc: TRUE
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/InToxSa_paper/Figure2_prepare"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we prepare the GWAS figure of the manuscript.

Design of the figure: left panel: phylogenetic tree right panel: multipanel with PCA of PI uptake, Lineage effects (+/- PI AUC distribution per CC) and Manhattan plot

could also add scatter plot with phenotypic and genetic distance

```{r message=FALSE}
library(tidyverse)
library(patchwork)
library(cols4all)
rm(list = ls())
```

# Import tree

```{r}
p1 <- readRDS("../Phylogeny/processed_data/tree_with_metadata.Rda")
```

# Import scatter plot

```{r}
rf_plots <- readRDS("../PCA/processed_data/RF_plots_with_CONTROLS.Rda")
p2 <- rf_plots[[4]]

p2 <- p2 +
  labs(x = "PI uptake AUC", y = "Maximum PI uptake rate", colour = "Toxicity cluster") +
  theme_bw(base_size = 11) +
  theme(text = element_text(face = "bold"),
        legend.position = c(0.3,0.7),
        legend.background = element_blank()) 
p2
```

# Import variable importance plot

```{r}
p3 <- rf_plots[[2]]

# labels <- p3$data$Variable %>%
#   str_match("(.*)_death_(mean|sd)") %>%
#   set_colnames(c("variable", "parameter", "statistic")) %>%
#   as_tibble() %>%
#   mutate(parameter_long = fct_recode(parameter,
#     `PI uptake AUC` = "AUC",
#     `Maximum PI uptake rate` = "max_rate",
#     `Peak PI uptake` = "max",
#     `Time of peak PI uptake` = "time_of_max",
#     `Time of maximum PI uptake rate` = "time_of_max_rate",
#     `Trough PI uptake`= "min",
#     `Time of through PI uptake` = "time_of_min"
#   )) %>%
#   mutate(parameter_short = fct_recode(parameter,
#     `AUC` = "AUC",
#     `Maximum rate` = "max_rate",
#     `Peak` = "max",
#     `Time of peak` = "time_of_max",
#     `Time of maximum rate` = "time_of_max_rate",
#     `Trough`= "min",
#     `Time of through` = "time_of_min",
#     `Final uptake` = "end_point"
#   )) %>%
#   mutate(label = str_c(parameter_short, " (", statistic, ")"))
# labels <- labels$label %>%
#   set_names(labels$variable)

p3 <- p3 +
  lims(x = c(20,80), y = c(0,.20)) +
  labs(x = "Gini index decrease", y = "Accuracy decrease") +
  theme_bw() +
  theme(text = element_text(face = "bold"))
```

# Import Lineage effects

```{r}
# p4 <- readRDS("../GWAS/Lineages/processed_data/lineage_effects_barplot.Rda")
p4 <- readRDS("../GWAS/Lineages/processed_data/lineages_rf_clusters_scatterplot.Rda") +
  ylim(-0.008,0.010) +
  labs(x = "MDS4", y = "MDS3", colour = "Toxicity cluster", fill ="Toxicity cluster" ) +
  theme(text = element_text(face = "bold"),
        legend.position = c(0.25,0.8),
        legend.background = element_blank())
p4 
```

# Import Manhattan plot

```{r}
pyseer_plots <- readRDS("../GWAS/pyseer_output/processed_data/pyseer_output_toxicity_genes_all_plots.Rda")
p5 <- pyseer_plots[[7]]
p5 <- p5 +
  labs(y = "-lgo10(p)") +
  theme(legend.position = c(.3,.7),
        legend.spacing.y = unit(1, units = "points"))
p5
```

# Compose final figure

```{r}
p6 <- wrap_plots(p3, p2, p4, p5, ncol = 2)

p1 + p6 +
  plot_layout(widths = c(4,4))+
  plot_annotation(tag_levels = "A")

# ggsave("figures/Figure2.pdf", width = 16, height = 8)
```

Try another

```{r}
p4 <- p4 + 
  labs(x = "", y = "-log(p)") +
  theme(panel.grid = element_blank(),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10))
p5 + patchwork::inset_element(p4, left = 0.05, bottom = .55, right = .5, top = .95)
```

# Import heritability

Calculate heritability for each PI parameter? (later)?

