---
title: "Raw data processing"
author: "R Guerillot"
date: "11/12/2020"
output: html_document
---

```{r setup, include=T, message=FALSE}
library(tidyverse)
library(here)
library(patchwork)
setwd(paste0(here(), "/5-Data_processing"))
getwd()
```

# Source all function

```{r message=FALSE}
rm(list = ls())
source("../0-Functions/all_functions.R")
```

# Import all VANANZ PI and growth curve data (plate 1-6 + P1-2, M1- GP1-7)

```{r}
raw_PI.df <- list.files(path="../3-Data_parsing/parsed_experiments/", pattern="PI", full.names=TRUE, recursive=T) %>% 
  map_df(~read_csv(., col_types = cols(well_row = col_factor(),
                                      well_col = col_factor(),
                                      f_signal = col_integer(),
                                      well = col_factor(),
                                      experiment = col_factor(),
                                      cell_number = col_factor(),
                                      plate = col_factor(),
                                      well_id = col_factor(),
                                      timepoint = col_integer(),
                                      sample_id = col_factor(),
                                      replicate = col_factor(),
                                      strain_group = col_factor()))) %>%
  filter(cell_number == 40000) %>%                        # Keep only data with 40000 HeLa cells
  mutate(experiment_id = str_c(experiment, "_", plate))   # Create a experiment variable that is unique
                                                          # -> two plate were done the same day => same exp.)

```

# Check plates and replicates

```{r}
print(paste("Total number of PI plates:", 
            raw_PI.df %>% select(experiment_id) %>% distinct() %>% nrow()))

nb_replicate_plate.df <- raw_PI.df %>% select(experiment, plate) %>% distinct() %>% count(plate,name = "number of plate replicate") %>% print()

sum(nb_replicate_plate.df$`number of plate replicate`)

```

nb experiment PI kinetics with 40.000 cellules = 40

# Check timepoints interval
```{r}
for (exp in raw_PI.df$experiment_id %>% unique()) {
  print(paste(exp,
               raw_PI.df %>%
                 filter(experiment_id == exp) %>%
                 .$timepoint %>%
                 unique() %>%
                 sort() %>%
                 .[[2]]
  )
  )
  
}
```

# Process PI kinetics data
## Check JE2 reference outlier in each plate

Use stringent criteria to detect JE2 outlier curves
 
Timepoint is outlier if:
  1. |x| > 1.5 x IQR (Tukey's fences method = box plot outlier detection method)
  2. sd > 0.2 (Only timepoints with important variation are considered; sd > 20% of JE2 max)
Curve is outlier if: 
  1. > 5% of timepoints are outliers (outlier_tukey_proportion = 0.05)

Notes: 
  - 3 x sd and 3 x mad methods don't identify more oulier curves as 0.05 proportion threshold; tukey is the most stringent method and therefore the only method used here 
  - sd > 0.2 has been empirically set to avoid flagging outliers with very little variations. Could be improved by defining the sample_id~timepoints wise sd across all dataset
  
```{r}
raw_PI_check <- raw_PI.df %>%
  check_reference(reference = "JE2",
                  outlier_tukey_proportion = 0.05, # 1.5 x IQR used
                  outlier_sd_proportion = 1,       # 3 x sd not used
                  outlier_mad_proportion = 1,      # no 3 x mad not used 
                  k_tukey = 1.5,
                  min_sd = 0.2,
                  filter = F,
                  plot = T)
```

### Plot JE2 non outliers and oultiers only
```{r}
raw_PI_check %>% 
  filter(outlier_curve == F) %>%
  plot_cell_death(fitted = T, all_replicates = T) +
  ylim(-1,2.5) 

raw_PI_check %>% 
  filter(outlier_curve == T) %>%
  plot_cell_death(fitted = T, all_replicates = T) +
  ylim(-1,2.5) 
```

### Count number of JE2 outliers per experiment
```{r}
JE2_outliers_exp_count <- raw_PI_check %>% 
  filter(sample_id == "JE2") %>% 
  select(experiment_id, well_id, outlier_curve) %>% 
  distinct(.) %>% 
  group_by(experiment_id) %>% 
  mutate(total_JE2 = n()) %>% 
  ungroup() %>%
  group_by(experiment_id, outlier_curve) %>%
  mutate(number_curve = n()) %>% 
  ungroup() %>%
  select(-well_id) %>%
  distinct() %>%
  filter(outlier_curve == T) %>%
  mutate(remaining_non_JE2_outliers = total_JE2 - number_curve) %>% 
  select(experiment_id, outlier_curve, number_curve, total_JE2, remaining_non_JE2_outliers) %>%
  print()
```

### Identifiy all well_id corresponding to JE2 ouliers from dataset
```{r}
JE2_outliers_well <- raw_PI_check %>% 
  filter(outlier_curve == T) %>%
  .$well_id %>%
  unique() %>%
  print()
```

### Identify all experiment_id with < 2 JE2 non outlier 
```{r}
JE2_outliers_exp <- JE2_outliers_exp_count %>% 
  filter(remaining_non_JE2_outliers < 2) %>%
  print() %>%
  .$experiment_id 
```

Based on JE2 outlier detection:
- exp191121_4 will be excluded: 2 JE2 among 3 are outliers 
- exp200313_M2 will be excluded: 2 JE2 among 3 are outliers
- exp200817_GP4 will be excluded: 4 JE2 among 4 are outliers
- exp200919_GP7 will be excluded: 7 JE2 among 8 are outliers
- exp201103_P2 will be excluded: 2 JE2 among 3 are outliers
- exp201116_M2 will be excluded: 2 JE2 among 3 are outliers
- exp201122_4 will be excluded: 2 JE2 among 3 are outliers

### Check how many plate with < 3 JE3 non-oulier 
```{r}
JE2_outliers_exp_count %>% 
  filter(remaining_non_JE2_outliers < 3) 
```

If we want at least 3 non outlier JE2 per plates the folowing plates should be excluded too:
- exp191115_1
- exp191120_4
- exp200219_P1
- exp200228_M1

## Clean raw PI data: 
1. remove JE2 well outliers
2. remove experiments with < 2 JE2
3. standardise (with max time cutting)
4. fit curves
5. identify and remove outliers (default parameter with tukey + mad + sd proportion > 0.1 (>10% outlier timepoints))
6. calculate derivatives

```{r}
clean_PI_with_outliers.df <- raw_PI.df %>%
  filter(!well_id %in% JE2_outliers_well) %>%
  filter(!experiment_id %in% JE2_outliers_exp) %>%
  standardise_curves(cut = T) %>%
  fit_curves(use_standardised = T) %>%
  flag_outlier_curves(use_standardised = T, 
                      min_sd = 0.2,
                      k_sd = 3, 
                      k_mad = 3,
                      k_tukey = 1.5, 
                      outlier_tukey_proportion = 0.1, 
                      outlier_sd_proportion = 0.1, 
                      outlier_mad_proportion = 0.1,
                      filter = F) %>%
  get_derivatives(use_fitted = T) 

clean_PI.df <- clean_PI_with_outliers.df %>%
  filter(outlier_curve == F)
```


### Count clean plates remaining
```{r}
print(paste("Total number of validated PI plates:", 
            clean_PI.df %>% select(experiment_id) %>% distinct() %>% nrow()))

clean_PI_plate_count.df <- clean_PI.df %>% select(experiment, plate) %>% distinct() %>% count(plate,name = "number of plate replicate") %>% print()    
```

### Count proportion of outlier wells per plate
```{r}
ratio_outlier.df <- clean_PI_with_outliers.df %>%
  select(well_id, experiment_id, outlier_curve) %>%
  distinct() %>%
  mutate(outlier_curve = factor(outlier_curve, levels = c(T,F))) %>%
  complete(outlier_curve) %>%
  count(experiment_id, outlier_curve, .drop = F) %>%
  group_by(experiment_id) %>%
  summarise(outlier = paste(outlier_curve, collapse = "/"), ratio_outlier = paste(n, collapse = "/")) %>%
  ungroup() %>%
  separate(col = ratio_outlier, sep = "/", into = c("outlier", "non_outlier")) %>%
  mutate(proportion_outlier = as.numeric(outlier)/as.numeric(non_outlier)) %>%
  arrange(-proportion_outlier) %>%
  print(n = 40)

```

### Identify excluded isolate sample (outliers)
```{r}
excluded_sample <- setdiff(
raw_PI.df %>% .$sample_id%>% unique(),
clean_PI.df %>% .$sample_id%>% unique()
)

raw_PI.df %>% filter(sample_id %in% excluded_sample) %>%
  select(experiment_id, sample_id) %>%
  distinct() 
```

### Asign experiment replicate number
```{r}
clean_PI.df <- clean_PI.df %>%
  select(plate, experiment_id) %>%
  distinct() %>%
  mutate(number = 1) %>%
  group_by(plate) %>%
  mutate(plate_replicate = as.factor(cumsum(number))) %>%
  ungroup() %>%
  select(experiment_id, plate_replicate) %>%
  merge(clean_PI.df, ., "experiment_id") 
```

### Set sample_id level to get controls first when plotting
```{r}
clean_PI.df <- clean_PI.df %>%
  mutate(sample_id = factor(sample_id, levels = c("JE2", "NE1354", "TOX-4", "NEB agrA", "Non infected", "Complete lysis", sort(unique(grep("BPH", as.character(clean_PI.df$sample_id), value = T))), "Blank")))

```

## Extract PI curves parameters
```{r}
clean_parameters_PI.df <- clean_PI.df %>%
  get_parameters()
```

## Calculate sample parameters mean, median, sd, mad, IQR
```{r}
clean_sample_parameters_PI.df <- clean_parameters_PI.df %>%
  group_by(sample_id) %>%
  select(sample_id, ends_with("_death")) %>%
  summarise_if(is.numeric, funs(mean, sd, median, mad, IQR)) %>%
  merge(., clean_parameters_PI.df %>% select(sample_id, -ends_with("_death")) %>% distinct(), by = "sample_id") 
```

# Save processed data

## Create file structure

```{r}
dir.create("processed_data/clinical_strains/")
dir.create("processed_data/clinical_strains/PI_kinetics")
dir.create("processed_data/clinical_strains/PI_parameters")
```

## Save processed clinical strain PI data
```{r}
# remove empty columns with only NA
clean_PI.df <- clean_PI.df[,colSums(is.na(clean_PI.df))<nrow(clean_PI.df)] %>%
  # keep only clinical strains in saved file,
  filter(strain_group != "NEBRASKA")

# save file
saveRDS(clean_PI.df, 
        file="processed_data/clinical_strains/PI_kinetics/PI_kinetics.Rds", 
        compress = "gzip")
```

## Save PI curves parameters
```{r}
# remove empty columns with only NA
clean_parameters_PI.df <- clean_parameters_PI.df[,colSums(is.na(clean_parameters_PI.df))<nrow(clean_parameters_PI.df)] %>%
  # keep only clinical strains in saved file,
  filter(strain_group != "NEBRASKA")

# save file
saveRDS(clean_parameters_PI.df,
          file="processed_data/clinical_strains/PI_parameters/PI_parameters.Rds")          
```