---
title: "Prepare figure 3"
author: "Stefano Giulieri"
date: "16/06/2022"
output:
  html_document:
    toc: TRUE
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/InToxSa_paper/Figure3_prepare"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we prepare the figure summarising the genetic pairs analysis.

Design of the figure: 

Upper half: plot showing that the pairs have been chosen to maximise phenotypic distance and minimise genetic distance; phylogenetic tree or phylogenetic network or PCA of snippy output; adaptive network.

Lower half: plot with number of independent mutations (or p value for enrichment) +/- CC, mutations minimap +/- CC, PI uptake for NEB mutants, Operetta for NEB mutants

```{r message=FALSE}
library(tidyverse)
library(patchwork)
library(cols4all)
rm(list = ls())
```

# Import modified Vulcano plot

```{r}
plot_list <- readRDS("../Genetic_pairs/processed_data/PI_pairs_data/pairs_metadata_plots.Rda")
p1 <- plot_list[[2]] 
p1
```

# Import pairs plot

```{r}
p2 <- plot_list[[4]] +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = c(.7, .5))
p2
```

# Import convergence plots

```{r}
conver_plots <- readRDS("../Genetic_pairs/processed_data/gene_convergence/convergence_plots.Rda")
```

# Tree

```{r}
p3 <- readRDS("../Genetic_pairs/processed_data/gene_convergence/tree_with_mutations.Rda") +
  ylim(0, 80)
p3
```


# Compose figure

```{r}
# upper
p4 <- p1 + p2 + p3
p4

# lower 
p5 <- wrap_plots(conver_plots[c(1:3,6)], nrow = 1) 
p5

# final
p6 <- p4 / p5  +
  plot_annotation(tag_levels = "A")
p6



# ggsave("figures/Figure3.pdf", width = 18, height = 12)
```



