---
title: "VANANZ dynamic cytotoxicity"
author: "Stefano Giulieri"
date: "03/09/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(magrittr)
library(readxl)
library(kableExtra)
if(!require("growthcurve")) devtools::install_github("briandconnelly/growthcurve", build_vignettes = TRUE)
library(growthcurve)
grofit <- "https://cran.r-project.org/src/contrib/Archive/grofit/grofit_1.1.1-1.tar.gz"
if(!require("grofit")) install.packages(grofit, repos = NULL, type = "source")
library(grofit)
if(!require("corrplot")) install.packages("corrplot")
library(corrplot)

rm(list = ls())
```


## Import data
```{r}
file <- "Preliminary_analysis/Raw_data/exp_190831/PI_kinetics/190831 LD PI Assay VANANZ 24 hours.xlsx"
n_skip = 14 # set number of lines to skip in the Excel sheet
timepoint_interval <- 0.1
# generates vector of timepoints (in minutes) of fluorescence measurement. Here: every 6 minutes until 22 hours and 42 minutes
timepoint_max <- 22.7
timepoints <- seq(0, timepoint_max, timepoint_interval) * 60 
# generate vector of column names based on timepoints and wavelength
col_names_WL <- c(
  str_c("WL493_", timepoints),
  str_c("WL535_", timepoints)
)
col_names <- c(
  "well",
  "sample",
  col_names_WL
)
data <- read_excel(file, col_names = col_names, skip = 14) 

# kable(data) %>% 
#   kable_styling() %>%
#   scroll_box(width = "100%", height = "200px")
```

## Now we transform the data into a long format
```{r}
data_long <- data %>%
  gather(key = condition,
         value = f_signal,
         starts_with("WL")) %>%
  separate(condition, into = c("wave_length", "time"), remove = FALSE) %>%
  mutate(timepoint = as.numeric(time),
         well_row = str_extract(well, "[A-H]"),
         well_col = str_extract(well, "\\d{2}")) %>%
  select(-condition)
    
# data_long %>%
#   kable() %>%
#   kable_styling() %>%
#   scroll_box(width = "100%", height = "200px")

# if (!dir.exists("dataframes")){
#   dir.create("./dataframes")
# }

# write_csv("dataframes/dynamic_cytotoxicity_long_format.csv")
```

## import PI endpoint with max mortality
```{r}
data_long <- read_excel("Preliminary_analysis/Raw_data/exp_190831/PI_endpoints/190831 LD PI Assay VANANZ endpoint  plus fixed permeabilised last well.xlsx", 
                             col_types = c("text", "text", "numeric", "numeric"), 
                             skip = 13,
                             col_names = c("well", "sample", "f_signal_endpoint_493","f_signal_endpoint_535")) %>%
  merge(data_long, ., by = "well")

```

## And here it the plot of the raw data
```{r message=FALSE}
data_long %>%
  ggplot(aes(x = timepoint, y = f_signal, color = wave_length)) +
  geom_point(size = .5) +
  scale_colour_discrete(name = "Wave length", labels = c("493", "535")) +
  facet_grid(well_row ~ well_col) +
  labs(x = "Time (min)",
       y = "Fluorescence intensity") +
  theme_bw() 

# if (!dir.exists("figures")){
#   dir.create("./figures")
# }
# ggsave("figures/dynamic_cytoxicity_plate2_raw.pdf", width = 16, height = 16)
```

## Only plot 535 WL with PI at end point
```{r}
data_long %>%
  filter(wave_length == "WL535") %>%
  ggplot(aes(x = timepoint, y = f_signal, color = wave_length)) +
  geom_point(size = .5) +
  facet_grid(well_row ~ well_col) +
  geom_hline(data_long, mapping = aes(yintercept = f_signal_endpoint_535)) +
  labs(x = "Time (min)",
       y = "Fluorescence intensity") +
  theme_bw() 
```

# Extract zero and total cell death and calculate %death
```{r}
# total mortality defines as the mean of total lysis on non-infected  
total_mortality_W535 <- data_long %>%
  filter(wave_length == "WL535") %>%
  filter(well %in% c("H10","H11","H12")) %>%
  .$f_signal_endpoint_535 %>%
  unique() %>%
  mean()

# zero mortality defined as the minimum fluorescence signal of the plate to avoid negative %pct cell death 
zero_mortality_W535 <- data_long %>%
  filter(wave_length == "WL535") %>%
  .$f_signal %>%
  min()

data_long <- data_long %>%
  mutate(total_cell_death = total_mortality_W535) %>%
  mutate(zero_cell_death = zero_mortality_W535) %>%
  mutate(pct_death = (f_signal-zero_cell_death)/(total_cell_death-zero_cell_death))

data_long %>%
  filter(wave_length == "WL535") %>%
  ggplot(aes(x = timepoint, y = pct_death, color = wave_length)) +
  geom_point(size = .5) +
  scale_colour_discrete(name = "Wave length", labels = c("493", "535")) +
  facet_grid(well_row ~ well_col) +
  labs(x = "Time (min)",
       y = "% cell death") +
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  theme_bw() 
```

## We now merge with the sample ids so that we can group replicates and fit a spline model
```{r message=FALSE}
# Merge with well data
file <- "plate_info/well_info_n843.csv"
well_info <- read_csv(file) %>%
  mutate(sample_id = if_else(sample_id == "TOX-5",
                             "JE2", sample_id)) %>%
  filter(plate_number == 2)
df <- data_long %>%
  left_join(well_info) %>%
  filter(!is.na(f_signal))

# Plot 
df %>%
  ggplot(aes(x = timepoint, y = f_signal, color =  wave_length, group = wave_length)) +
  geom_point(size = .5) +
  scale_colour_discrete(name = "Wave length", labels = c("493", "535")) +
  facet_wrap(~sample_id) +
  labs(x = "Time (min)",
       y = "Fluorescence intensity") +
  growthcurve::stat_growthcurve(model = "spline", color = "black")+
  theme_bw()
# ggsave("figures/dynamic_cytoxicity_plate2.pdf", width = 16, height = 16)
```

## Are there difference between the two different wave length?
```{r fig.height=30, fig.width=10}
df %>%
  ggplot(aes(x = timepoint, y = f_signal, color =  wave_length)) +
  geom_point(size = .5) +
  scale_colour_discrete(name = "Wave length", labels = c("493", "535")) +
  facet_wrap(sample_id~wave_length, scales = "free_y", ncol = 4) +
  labs(x = "Time (min)",
       y = "Fluorstatescence intensity") +
  growthcurve::stat_growthcurve(model = "spline", color = "black")+
  theme_bw()
```

## Fitting a model to the curves with grofit and extracting curves parameters

```{r}
require(grofit)

# Remove initial cell death decrease for fitting
df <- df %>%
  filter(timepoint > 50)

# format time matrix for grofit
time_vector <- unique(df$timepoint)
len_matrix <- length(unique(df$well))
time.mat <- matrix(rep(time_vector,each=len_matrix),nrow=len_matrix)

# format data df for grofit
df_WL535 <- df %>% 
  filter(wave_length == "WL535") %>%
  select(well, timepoint, f_signal, sample_id) %>%
  reshape2::dcast(., well~timepoint, value.var = "f_signal") %>%
  mutate(concentration = 0.1, meta = "test") %>%
  select(well, meta, concentration, everything())

head(df_WL535)

# gofit growth curve fitting
res <- grofit::gcFit(time = time.mat,
                     data = df_WL535, 
                     grofit.control(fit.opt = "s",    # spline fit
                                    smooth.gc = 0.8,  # smoothness parameter
                                    interactive = F)) # non interactive

# extract/rename fitted parameters
res.df<-res$gcTable %>%
  rename(well = TestId) %>%
  select(well, mu.spline, A.spline, integral.spline) %>% # lambda parameter<=>lag_phase not fitted correctly
  rename(AUC = integral.spline, max_tox_rate = mu.spline, max_tox = A.spline)

# merge results with data and sort
df_new <- merge(df, res.df, by = "well", all.x = T)
df_new <- transform(df_new, well=reorder(well, -AUC))  # can change sorting variable here 

# plot sorted data
ggplot(df_new %>% filter(wave_length == "WL535"), aes(x = timepoint, y = f_signal, color = wave_length)) +
  geom_point(size = .5) +
  growthcurve::stat_growthcurve(model = "spline", color = "black")+
  facet_wrap(~sample_id ) +
  labs(x = "Time (min)",
       y = "Fluorescence intensity") +
  theme_bw()
```

## Create a dataframe with summarised toxicity parameters
```{r}
df_summary <- df_new %>%
  filter(time == max(time) & wave_length == "WL535") %>%
  group_by(well) %>%
  mutate(max_PI = max(f_signal)) %>%                                    # maximum PI value from raw data
  ungroup() %>%
  group_by(sample_id) %>%
  summarise(mean_pct_death = mean(pct_death),
            sd_pct_death = sd(pct_death),
            mean_max_tox = mean(max_tox),                               # maximum PI value from fitted curve
            sd_max_tox = sd(max_tox),
            mean_max_tox_rate = mean(max_tox_rate),
            sd_max_tox_rate = sd(max_tox_rate),
            mean_AUC = mean(AUC),
            sd_AUC = sd(AUC),
            mean_max_PI = mean(max_PI),
            sd_max_PI = sd(max_PI),
            mean_f_signal_endpoint_535 = mean(f_signal_endpoint_535),
            sd_f_signal_endpoint_535 = sd(f_signal_endpoint_535)) %>%
  ungroup()
```

## import/summarise LDH data and merge with PI results summary
```{r}
LDH.df <- read_excel("../Raw_data/exp_190831/LDH/190831 plate 2 VANANZ HeLa cells infected 24h post PI cell death assay.xlsx",
                     col_types = c("text", "numeric", "text", "numeric", "numeric", "numeric"),
                     col_names = c("row", "col", "content", "raw_LDH", "blank_cor_LDH", "blank_cor_LDH2"),
                     skip = 13)

# Create well variable and permit merge with plateinfo
require(stringr)
LDH.df <- LDH.df %>%
  mutate(col = stringr::str_pad(col, 2, pad = "0")) %>%
  mutate(well = paste0(row, col)) %>%
  merge(well_info, ., by = "well")

# Summarise LDH results
LDH.summary <- LDH.df %>%
  group_by(sample_id) %>%
  summarise(mean_LDH = mean(blank_cor_LDH),
            sd_LDH = sd(blank_cor_LDH))%>%
  ungroup()%>%
  mutate(mean_LDH = ifelse(is.na(mean_LDH), 0, mean_LDH)) %>% # replace NA value by 0 for blank
  mutate(sd_LDH = ifelse(is.na(sd_LDH), 0, sd_LDH)) # replace NA value by 0 for blank

# Merge PI and LDH result summaries

df_summary <- merge(df_summary, LDH.summary, by = "sample_id")

```

## Correlation analysis of the parameters

```{r}
# transform to matrix for cor function
rownames(df_summary) <- df_summary$sample_id
df_summary2 <- df_summary %>%
  select(-sample_id) %>%
  select_if(grepl("mean", names(.)))

cor.mat = cor(df_summary2, method = "pearson") # spearman rank correlation
cor.df<-as.data.frame(cor.mat)

# function to compute a correlation significance matrix 
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(method = "pearson", mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

# compute significance of correlation matrix
cor.signif <- cor.mtest(cor.mat)
cor.signif

# visualise correlation matrix
require(corrplot)
corrplot(cor.mat,
         #p.mat = cor.signif,
         sig.level = 0.05,
         addCoef.col = "white",
         number.cex = .6,
         order="hclust", 
         tl.col="black")

plot(df_summary$mean_LDH, df_summary$mean_f_signal_endpoint_535)

```