---
title: "Get correlations between tree branches and PC"
author: "Stefano Giulieri"
date: "01/11/2022"
output: html_document
---


```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/InToxSa_paper/GWAS/Lineages"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we get correlations between tree branches and principal components. This will allow us to map the dominant PC on tree branches as in Earle et al, Nat Microbiol 2016.

This analysed is based on the `bugwas` code reported here: https://github.com/sgearle/bugwas/blob/master/bugwas/R/BUGWAS_functions.R#L962

Idea on how to improve this: infer ancestral state for each PC and map on tree (either PC with highest value or a tree for each PC).

```{r}
library(tidyverse)
library(ggtree)
library(patchwork)
library(ggridges)

rm(list = ls())
```

# Import data

Phylogenetic tree

```{r}
tree <- read.tree("../../Phylogeny/raw_data/core90.aln.treefile")
tree <- tree %>%
  phytools::midpoint.root() %>%
  ape::drop.tip("Reference")
tree$edge.length <- tree$edge.length*3.0496e6

tree
```

Principal components

```{r}
pca <- readRDS("processed_data/MDS/df_mds_mash.Rda")
pca <- pca %>%
  column_to_rownames("sample_id") %>%
  as.matrix()

str(pca)
```

# Get correlations

Get tree patterns

```{r}
tree2patterns <- function(tree = NULL,
						  tiporder = NULL) {
						  	
	if(is.null(tiporder)){
		tiporder <- tree$tip.label
	}
	n <- length(tree$tip.label)
	
	
	mtp <- matrix(0,n,n+tree$Nnode)
	
	#Sorting out tree patterns for external nodes.
	mtp[,1:n] <- diag(n)
	
	#Sorting out tree patterns for internal nodes.
	for(i in 1:tree$Nnode) {
		wh <- match(ape::extract.clade(tree, n+i)$tip.label, tree$tip.label)
		mtp[wh, n+i] <- 1
	}
	
	#Reorder the rows to match the order of the individuals
	mtp <- mtp[match(tiporder, tree$tip.label), ]
	
	if(is.null(tree$node.label)){
		tree$node.label = paste("node",1:tree$Nnode+n, sep="")
	}
	
	
	
	mtp.f <- apply(mtp, 2, mean)
	mtp[, mtp.f>0.5] <- 1-mtp[, mtp.f>0.5]
	mtp.f <- apply(mtp, 2, mean)
	
	#Label the rows and columns of tree patterns.
	rownames(mtp) <- tiporder
	colnames(mtp) = c(tiporder, tree$node.label)
	
	
	edge <- match(1:ncol(mtp), tree$edge[, 2])
	edge.length <- tree$edge.length[edge]
	return(list("pat" = mtp, "labelled.tree" = tree,
				"ancestral_edge" = edge,
				"ancestral_edge.length" = edge.length))
}
tree_pat <- tree2patterns(tree)

str(tree_pat)

mtp <- tree_pat$pat

mtp[-1,]

# heatmap(mtp)
```

Get correlations

```{r}
get_correlations <- function (XX = NULL, 
							  pca = NULL,
							  npcs = NULL,
							  id = NULL,
							  all.cor  = FALSE){
	
	
	cor.XX.pca <- cor(XX,pca[, 1:npcs])
	cor.XX.pca[is.na(cor.XX.pca)] = 0
	which.pc <- apply(abs(cor.XX.pca), 1, which.max)
	max.cor.pc <- apply(abs(cor.XX.pca), 1, max)
	if(all.cor){
		return(list("which.pc" = which.pc, "max.cor.pc" = max.cor.pc, "all.cor.pc" = cor.XX.pca))
	}else{
		return(list("which.pc" = which.pc, "max.cor.pc" = max.cor.pc))
	}
	
}

cor_tree <- get_correlations(
		XX = mtp, pca = pca, npcs = ncol(pca), id = tree$tip.label, all.cor = TRUE)
str(cor_tree)

df_cor_tree  <- tibble(label = names(cor_tree$which.pc),
                       PC = cor_tree$which.pc,
                       PC2 = if_else(PC <=10, PC, NA_integer_),
                       cor = cor_tree$max.cor.pc) %>%
  bind_cols(as_tibble(cor_tree$all.cor.pc))

```

Plot PCs

```{r}
ggtree(tree_pat$labelled.tree) %>%
  .$data %>%
  filter(!isTip)

ggtree(tree_pat$labelled.tree) %<+% (df_cor_tree %>% filter(str_detect(label, "BPH", negate = T))) +
  geom_tree(aes(colour = as.factor(PC2))) +
 scale_colour_viridis_d() +
  geom_nodelab(aes(label = PC2))
```


Plot correlations with single PC

```{r}
plot_list <- list()
gg <- ggtree(tree_pat$labelled.tree) %<+% df_cor_tree 
for (pc in colnames(cor_tree$all.cor.pc)){
  s <- rlang::sym(pc)
  p <- gg +
  geom_tree(aes(colour = {{s}})) +
 scale_colour_gradient2()
  
  plot_list[[pc]] <- p
}

wrap_plots(plot_list)
```

```{r}
plot_list <- list()
gg <- ggtree(tree_pat$labelled.tree) %<+% df_cor_tree 
tree_rescaled <- tidytree::as.treedata(gg) %>%
  as_tibble() %>%
  mutate(rescaled = sqrt(branch.length)) 

tree_rescaled <- treeio::rescale_tree(tidytree::as.treedata(tree_rescaled), branch.length = "rescaled")
(gg <- ggtree(tree_rescaled, layout = "rectangular"))
# (gg <- ggtree(tree_rescaled, layout = "circular"))



for (pc in colnames(cor_tree$all.cor.pc)){
  s <- rlang::sym(pc)
  p <- gg +
  geom_tree(aes(colour = {{s}}), size = 1) +
 scale_colour_gradient2(limits = c(-.4,.4), name = "Correlation") +
    labs(title = pc) +
    theme(title = element_text(size = 16),
          legend.text = element_text(size = 14))
  
  plot_list[[pc]] <- p
}

plot_list[[1]]

p <- wrap_plots(plot_list, nrow = 3) +
  plot_layout(guides = "collect")

# ggsave("figures/tree_with_pc_correlations.png", p, width = 16, height = 12)
```

Set correlation threshold

```{r}
gg$data %>%
  ggplot(aes(x = cor, fill = isTip, colour = isTip)) +
  geom_density(alpha = .5) +
  theme_bw()

gg$data %>%
  ggplot(aes(x = x, y = cor)) +
  geom_point() +
  facet_wrap(~isTip, ncol = 1) +
  theme_bw()
```

A threshold of .2 seems reasonable

```{r}
# tree_cor<- tidytree::as.treedata(gg) %>%
#   as_tibble() %>%
#   mutate(PC2 = if_else(cor > .2, PC, NA_integer_)) %>%
#   treeio::as.treedata()
gg +
  geom_tree(aes(colour = as.factor(PC))) +
 scale_colour_viridis_d()
```

Update annotated tree (later)

```{r}
# p <- readRDS("processed_data/tree_with_metadata_ALL.Rda")
```

# Get correlations with major ST

```{r}
tree_data <- readRDS("../../Phylogeny/processed_data/curated_metadata_tree.Rda")

top_st <- tree_data %>%
  filter(ST != "-") %>%
  count(ST, sort = T) %>%
  # filter(n >= 20) %>%
  slice_head(n = 15) %>%
  .$ST %>%
  as.character()

st <- tree_data %>%
  filter(ST %in% top_st) %>%
  select(sample_id, ST) %>%
  mutate(value = 1) %>%
  pivot_wider(names_from = "ST", values_from = "value", values_fill = 0) %>%
  column_to_rownames("sample_id") %>%
  as.matrix()

str(st)

pca <- pca[match(rownames(st), rownames(pca)),]

cor_st <- get_correlations(XX = st, pca = pca, npcs = 10, all.cor = TRUE)

df_cor_st <- as_tibble(cor_st$all.cor.pc, rownames = "ST")

cor_st$all.cor.pc %>%
  heatmap()

p1 <- df_cor_st %>%
  pivot_longer(cols = starts_with("MDS"), names_to = "PC", values_to = "cor") %>%
  mutate(PC_num = as.integer(str_remove(PC, "MDS"))) %>%
  ggplot(aes(x = fct_rev(fct_relevel(ST, top_st)), y = fct_reorder(PC, PC_num), fill = cor)) +
  geom_tile() +
  scale_fill_gradient2(name = "Correlation") +
  labs(x = "Sequence type", y = "Principal component") +
  coord_flip() +
  theme_bw(base_size = 16) +
  theme(text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))

 p2 <- tree_data %>%
   filter(ST %in% top_st) %>%
  ggplot(aes(x = pi_auc, y = fct_rev(fct_relevel(ST, top_st))))+ 
  geom_density_ridges() +
   labs(y = "", x = "PI uptake AUC") +
  theme_bw(base_size = 16) +
   theme(
         text = element_text(face = "bold"))
 
 p1 + p2 +
   plot_annotation(tag_levels = "A")
 
 # ggsave("figures/cc_major_pc_heatmap_pi_auc_ridges.pdf")
 # 
 # ggsave("figures/cc_major_pc_heatmap.pdf", p1)
```

