---
title: "Perform PCA to infer lineages (vcf of mutations)"
author: "Stefano Giulieri"
date: "24/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c("~/Documents/Github/InToxSa", "/6-InToxSa_paper/GWAS/Lineages"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we perform multi-dimensional scaling (MDS) to infer major lineages in the VANANZ PI dataset (387 strains). We then use this lineage information to infer impact on heritability and for the ML model.

```{r}
library(tidyverse)

rm(list = ls())
```

# Multi-dimensional scaling using mash distances

Here we use multidimensional scaling from the mash distance matrix. We use the `cmdscale` function from base R. A more modern approach using the `recipes` and the `step_isomap()` function is described here:
https://recipes.tidymodels.org/reference/step_isomap.html
https://www.tidyverse.org/blog/2021/03/workflowsets-0-0-1/
https://juliasilge.com/blog/cocktail-recipes-umap/

```{r}
f <- "~/Documents/Github/Transfer_with_server/mash.dist.tab"
dir <- "raw_data/"
file.copy(f, dir)
mash <- as.matrix(read.delim(str_c(dir, basename(f)),as.is=T,h=T,sep="\t",comment.char='',check.names=FALSE, row.names = 1))
str(mash)
loc2 <- cmdscale(mash, k = 2, eig = T)
plot(loc2$eig[1:30])

df_eigen <- tibble(eigenvalue = loc2$eig, 
       PC = seq_along(loc2$eig)) %>%
  # remove negative eigenvalues as in https://github.com/mgalardini/pyseer/blob/master/pyseer/cmdscale.py
  filter(eigenvalue > 0) %>% 
  mutate(variance_explained = eigenvalue/sum(eigenvalue))

df_eigen %>%
  slice_head(n = 30) %>%
  ggplot(aes(x =PC, y = eigenvalue)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(n.breaks = 30) +
  theme_bw()

df_eigen %>%
  slice_head(n = 30) %>%
  ggplot(aes(x =PC, y = variance_explained)) +
  geom_line() +
  scale_x_continuous(n.breaks = 30) +
  theme_bw()

df_eigen %>%
  mutate(cum_var = cumsum(variance_explained)) %>%
  slice_head(n = 30) %>%
  ggplot(aes(x = PC, y = cum_var)) +
  geom_line() +
  scale_x_continuous(n.breaks = 30) +
  theme_bw()

# repeat using the first 5 dimensions
loc3 <- cmdscale(mash, k = 5, eig = F)
x <- loc3[, 1]; y <- loc3[, 2]
plot(x, y, xlab = "", ylab = "",
    main = "Multidimensional Scaling Analysis (IBS)")

# repeat using the first 10 dimensions
loc4 <- cmdscale(mash, k = 10, eig = F)
x <- loc4[, 1]; y <- loc4[, 2]
plot(x, y, xlab = "", ylab = "",
    main = "Multidimensional Scaling Analysis (IBS)")
```


```{r}
colnames(loc4) <- str_c("MDS", 1:ncol(loc4))
df_mds_mash <- as_tibble(loc4, rownames = "sample_id")
```


Plot the principal component pairs for the first five PCs:

Here is a good examples
https://juliasilge.com/blog/billboard-100/


```{r}
# df_mds_mash %>%
#   select(-sample_id) %>%
#   select(num_range("MDS", 1:5)) %>%
#   GGally::ggpairs() +
#   theme_bw()

# df_mds_mash %>%
#   select(-sample_id) %>%
#   select(num_range("MDS", 1:5)) %>%
#   GGally::ggpairs(lower = list(continuous = "points", combo = "facethist", discrete = "facetbar", na =
#     "na")) +
#   theme_bw()

df_mds_mash %>%
  select(-sample_id) %>%
  select(num_range("MDS", 1:5)) %>%
  ggplot() +
  ggforce::geom_autopoint() +
  ggforce::geom_autodensity(alpha = .3) +
  ggforce::facet_matrix(vars(everything()), 
                        layer.diag = 2) +
  theme_bw()

df_mds_mash %>%
  select(-sample_id) %>%
  select(num_range("MDS", 1:5)) %>%
  ggplot() +
  ggforce::geom_autopoint() +
  ggforce::geom_autodensity(alpha = .3) +
  geom_density2d(aes(x = .panel_x, y = .panel_y)) +
  ggforce::facet_matrix(vars(num_range("MDS", 1:5)),
                        layer.lower = 1,
                        layer.diag = 2,
                        layer.upper = 3) +
  theme_bw()
```

# Save processed data

```{r}
dir <- "processed_data/MDS/"
dir.create(dir)

df_mds_mash %>%
  saveRDS(str_c(dir, "df_mds_mash.Rda"))

```

# Map RF clusters from the proximity matrix

```{r}
df_rf <- readRDS("../../PCA/processed_data/df_RF_clusters_with_CONTROLS.Rda")
```

Map on the 2 components with largest lineage effect

```{r}
df_plot <- df_mds_mash %>%
  inner_join(df_rf)

(p1 <- df_plot %>%
  ggplot(aes(x = MDS4, y = MDS3, colour = cluster)) +
  geom_point() +
  stat_ellipse(geom = "polygon", aes(fill = cluster), alpha = .3) +
  scale_color_manual(values = c("orange", "red", "blue")) +
   scale_fill_manual(values = c("orange", "red", "blue")) +
  theme_bw())


p1 %>%
  saveRDS("processed_data/lineages_rf_clusters_scatterplot.Rda")

```

Explore correlations

```{r}
df_plot %>%
  select(-sample_id) %>%
  select(num_range("MDS", 1:5), AUC_death_mean) %>%
  ggplot() +
  ggforce::geom_autopoint() +
  ggforce::geom_autodensity(alpha = .3) +
  ggforce::facet_matrix(vars(everything()), 
                        layer.diag = 2) +
  theme_bw()
```

