---
title: "ML phylogenetic tree of all VANANZ strains with PI phenotype available"
author: "Stefano Giulieri"
date: "24/05/2022"
output:
  html_document:
    toc: TRUE
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/InToxSa_paper/Phylogeny"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we visualise the ML tree of all VANANZ strains with PI phenotype (strains included in the GWAS, n = 387).

```{r message=FALSE}
library(tidyverse)
library(skimr)
library(ggtree)
library(phytools)
library(ggnewscale)
rm(list = ls())
```

# Load raw data: tree and metadata

## Tree in Newick format

Here we load the initial tree. ***Please update when iqtree has finished***

```{r}
tree <- read.tree("processed_data/clean.core90.aln.tree")
```

## Metadata

PI uptake AUC

```{r}
pheno <- readRDS("../GWAS/Phenotypes/processed_data/toxicity_phenotype_with_norm.Rda") %>%
  select(sample_id, pi_auc = untransformed)
```

sequence type and mecA

```{r}
sequence_data <- read_csv("../../plate_info/strain_metadata.csv") %>%
  transmute(sample_id,
            ST = as.factor(ST),
            mecA = factor(if_else(mecA == 1, "present", "absent")))
```

inferred CC

```{r}
cc_data <- readRDS("processed_data/cc_phylo_inferred.Rda")
```

Contrasting PI pairs

```{r}
PI_pairs <- read_csv("../../Genetic_pairs_analysis_Oct_2020/processed_data/monophyletic_clusters/genetic_pairs_significant_PI_change.csv") %>%
  pivot_longer(cols = dplyr::everything(), names_to = "iso_class", values_to = "sample_id") %>%
  mutate(iso_class = as.factor(iso_class))
```


This is the metadata file

```{r}
df_metadata <- pheno %>%
  inner_join(sequence_data) %>%
  inner_join(cc_data) %>%
  full_join(PI_pairs)

skim(df_metadata)
```

# Plot the tree

## Basic

Rectangular tree

```{r}
gg_r <- ggtree(tree, layout = "rectangular") 
gg_r +
  geom_treescale(width = 1e4) +
  theme_bw()
```

Circular tree

```{r}
gg_c <- ggtree(tree, layout = "circular") 
x_scale <- 4e4
y_scale <- max(ggplot_build(gg_c)$layout$panel_params[[1]]$theta.range)/2
gg_c +
  geom_treescale(width = 1e4,
                 x = x_scale,
                 y = y_scale)  +
  theme_bw()
```

We prefer the rectangular tree (for now)

```{r}
gg <- gg_r
```

## Add major CC (>= 15 sequences)

```{r}
my_offset <- 1e3
my_fontsize <- 3
cc <- df_metadata %>%
  drop_na(CC) %>%
  count(CC, sort = T) %>%
  # filter(n >= 20) %>%
  slice_head(n = 15) %>%
  .$CC
p <- gg
for (my_cc in cc){
  my_node <- df_metadata %>% filter(CC == my_cc) %>% .$mrca %>% unique()
  p <- p +
    geom_cladelab(node = my_node,
                    label = str_c("CC", my_cc),
                    align = T,
                    offset = my_offset,
                    fontsize = my_fontsize,
                  fontface = "bold",
                    family = "Helvetica")
}
```

## Add mecA

```{r}
my_offset <-  .5e4
width <- .05
data <- df_metadata %>%
  select(sample_id, mecA) %>%
  column_to_rownames("sample_id")

p1 <- gheatmap(p, data, colnames = F, width = width, offset = my_offset, color = NA) +
  scale_fill_manual(values = c("white", "black"), label = c("MSSA", "MRSA"),name = "MRSA status", guide = guide_legend(order = 2) ,na.translate = F)
```

## Add PI uptake AUC

```{r}
width <- .05
# offset <- max(ggplot_build(p1)$layout$panel_params[[1]]$x.range) -
#   max(ggplot_build(p)$layout$panel_params[[1]]$x.range) 
offset <- .5e4 + width*max(ggplot_build(gg)$layout$panel_params[[1]]$x.range) 
data <- df_metadata %>%
  select(sample_id, pi_auc) %>%
  column_to_rownames("sample_id")
p2 <- p1 + new_scale_fill()
p2 <- gheatmap(p2, data, colnames = F, width = width, offset = offset, color = NA) +
  scale_fill_viridis_c(name = "PI uptake (AUC)", guide = guide_legend(order = 3), option = "plasma", na.value = "white")
```

Add PI pairs

```{r}
x <- max(ggplot_build(p2)$layout$panel_params[[1]]$x.range) + 1e3
p3 <- p2 + new_scale_fill()
p4 <- p3 %<+% df_metadata +
  geom_tippoint(aes(subset = !is.na(iso_class), fill = iso_class), shape = 21, x = 4.8e4) +
  scale_fill_manual(values = c("black", "white"), labels = c("reference strain (high toxicity)",
                                                             "paired strain (low toxicity)"),
                    name = "Pairs with contrasting PI phenotype") +
  theme(legend.position = c(.2,.7))
```


## Final changes

```{r}
p5 <- p4 +
  geom_treescale(width = 1e4,
                 x = .2,
                 y = .1) +
  theme(text = element_text(face = "bold"))
p5
```

# Save processed data

## Tree object

```{r}
processed_tree <- "tree_with_metadata"
dir <- "processed_data/"
saveRDS(p5, str_c(dir, processed_tree, ".Rda" ))
```

## Curated tree metadata

```{r}
df_metadata %>%
  saveRDS(str_c(dir, "curated_metadata_tree.Rda"))
```

## Tree figure

```{r}
dir <- "figures/"
dir.create(dir)
ggsave(str_c(dir, processed_tree, ".pdf"),
       width = 10, 
       height = 7)
```

TO DO: plot main lineages on the tree. We can look at the code here: 
https://github.com/sgearle/bugwas/blob/master/bugwas/R/BUGWAS_functions.R#L962



