---
title: "Generate GWAS output table"
author: "Stefano Giulieri"
date: "27/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(paste0(here::here(), "/6-InToxSa_paper/Generate_datasets"))
getwd()
```

Here we prepare supp table 4 with PI parameters. 3 excel sheets: clinical + controls, Neb + JE2 control, mutagenesis + BPH3370 control

```{r}
library(tidyverse)
library(writexl)
rm(list = ls())
```

# Assign table name

```{r}
table_name <- "supp_table_4"
```

# Import PI parameters: clinical isolates

```{r}
clinical <- readRDS("../../5-Data_processing/processed_data/clinical_strains/PI_parameters/PI_sample_parameters.Rda")
```

We keep clinical isolates (BPHXXXX) *and* controls. Also we reformat as tibble

```{r}
controls <- c("JE2", "TOX-4")
clinical <- clinical %>%
  as_tibble() %>%
  filter(str_detect(sample_id, "BPH") | sample_id %in% controls)
```

Check that there are no duplicates. Inspect the dataframe

```{r}
n_distinct(clinical$sample_id)
nrow(clinical)

skimr::skim(clinical)
```

Now we can reduce variables and add to spreadsheet list

```{r}
clinical <- clinical %>%
  select(sample_id, ends_with("mean"), ends_with("sd")) %>%
  select(-c(starts_with("end_point_timepoint_death"),
            starts_with("doubling_time")))

supp_table_prepare <- list(`clinical isolates` = clinical)
```

# Import PI parameters: Nebraska transposon mutants

```{r}
neb <- readRDS("../../5-Data_processing/processed_data/neb_mutants/PI_parameters/PI_sample_parameters_plateGPV1.Rda")

controls <- c("JE2", "TOX-4")
neb <- neb %>%
  as_tibble() %>%
  filter(str_detect(sample_id, "NE") | sample_id %in% controls)

n_distinct(neb$sample_id)
nrow(neb)

skimr::skim(neb)
```

Add minimal metadata

```{r}
neb_metadata <- read_csv("raw_data/df_neb_curated.csv") %>%
  select(sample_id = neb_mutant_id, unique_gene_symbol = neb_gene_symbol, neb_locus_tag)

neb <- neb %>%
  left_join(neb_metadata)
```

Reduce variables and add to spreadsheet list

```{r}
neb <- neb %>%
  select(sample_id, unique_gene_symbol, neb_locus_tag, ends_with("mean"), ends_with("sd")) %>%
   select(-c(starts_with("end_point_timepoint_death"),
            starts_with("doubling_time")))

supp_table_prepare$`Nebraska transposon mutants` <- neb
```

# Import PI parameters: allelic exchange 

```{r}
mutants <- readRDS("../../5-Data_processing/processed_data/allelic_exchange/PI_parameters/PI_sample_parameters_plateGP-C1.Rda") %>%
  as_tibble()

controls <- c("JE2", "TOX-4", "BPH3370")
mutants <- mutants %>%
  as_tibble() %>%
  filter(str_detect(sample_id, "BPH3370") | sample_id %in% controls)

n_distinct(mutants$sample_id)
nrow(mutants)
```

Add minimal metadata

```{r}
mutants_metadata <- read_csv("raw_data/allelic_exchange_metadata.csv") %>%
  select(sample_id, mutation = label) %>%
  filter(str_detect(sample_id, "BPH3370")) %>%
  mutate(mutation = if_else(sample_id == "BPH3370", "wildtype", mutation)) %>%
  distinct()

mutants <- mutants %>%
  left_join(mutants_metadata)
```

Reduce variables and add to spreadsheet list

```{r}
mutants <- mutants %>%
  select(sample_id, mutation, ends_with("mean"), ends_with("sd")) %>%
  select(-c(starts_with("end_point_timepoint_death"),
            starts_with("doubling_time")))

supp_table_prepare$`allelic exchange mutants` <- mutants
```

# List of variables 

```{r}
variables <- unique(c(colnames(clinical), 
                      colnames(neb), 
                      colnames(mutants)))
variables <- tibble(variable = variables,
                    comment = NA_character_)

supp_table_prepare$`variables` <- variables
```

# Export as supplementary table

```{r}
dir.create(str_c("processed_data/", table_name))

supp_table_prepare %>%
  write_xlsx(str_c("processed_data/", table_name, "/", table_name, ".xlsx"))
```
