---
title: "PCA of PI uptake data"
author: "Stefano Giulieri"
date: "25/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(paste0(here::here(), "/InToxSa_paper/PCA"))
getwd()
```

Here we generate repeat the PCA analysis of PI uptake data (already done by Romain). The idea is to show the rationale of chosing the AUC as summary measure for the phenotype

```{r}
library(tidyverse)
library(tidymodels)
library(skimr)
library(factoextra)
library(adegenet)
library(randomForest)
library(cluster)
library(caret)
library(patchwork)
library(corrr)
rm(list = ls())

source("../../Functions/all_functions.R")
```

# Import summary PI data

```{r}
PI_parameters <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/PI/dataframes/parameters/PI_sample_parameters.Rda")
```

We keep clinical isolates only (BPHXXXX). Also we reformat as tibble

```{r}
PI_parameters <- PI_parameters %>%
  as_tibble() %>%
  filter(str_detect(sample_id, "BPH"))
```

Check that there are no duplicates. Inspect the dataframe

```{r}
n_distinct(PI_parameters$sample_id)
nrow(PI_parameters)

skim(PI_parameters)
```

# Inspect/visualise/assess distributions

```{r}
vars <- sort(colnames(PI_parameters[-1]))
plot_list <- list()
for (v in vars) {
  s <- rlang::sym(v)
  
  p <- PI_parameters %>%
    ggplot(aes(x = {{s}})) +
    geom_histogram() +
    theme_bw()
  
  plot_list[[v]] <- p
}

wrap_plots(plot_list, ncol = 5)
ggsave("figures/all_vars_histograms.png", width = 12, height = 16)
```

Alternative representation

```{r}
PI_parameters %>%
  pivot_longer(cols = -sample_id, names_to = "parameter_stat", values_to = "value") %>%
  mutate(parameter =  str_match(parameter_stat, "(.*)_([^_]*)")[,2],
         statistic = str_match(parameter_stat, "(.*)_([^_]*)")[,3]) %>%
  select(-parameter_stat) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_grid(statistic~parameter, scales = "free")
```

Remove variable with no variance and doubling time (this is just a different representaiton of death rate)

```{r}
PI_parameters <- PI_parameters %>%
  select(-c(contains("end_point_timepoint"), 
            contains("doubling_time")))
```

Repeat histograms

```{r}
PI_parameters %>%
  pivot_longer(cols = -sample_id, names_to = "parameter_stat", values_to = "value") %>%
  mutate(parameter =  str_match(parameter_stat, "(.*)_([^_]*)")[,2],
         statistic = str_match(parameter_stat, "(.*)_([^_]*)")[,3]) %>%
  select(-parameter_stat) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_grid(statistic~parameter, scales = "free") +
  theme_bw()
ggsave("figures/all_vars_histograms_facet_grid.png")

PI_parameters %>%
  pivot_longer(cols = -sample_id, names_to = "parameter_stat", values_to = "value") %>%
  mutate(parameter =  str_match(parameter_stat, "(.*)_([^_]*)")[,2],
         statistic = str_match(parameter_stat, "(.*)_([^_]*)")[,3]) %>%
  select(-parameter_stat) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
   facet_wrap(parameter~statistic, scales = "free", ncol = 5) +
  theme_bw()
ggsave("figures/all_vars_histograms_facet_wrap.png", width = 9, height = 11)
```

# Assess correlations: all parameters

```{r}
corr <- PI_parameters %>%
  select(-c(sample_id)) %>%
  correlate() 

corr %>%
  rearrange() %>%
  shave() %>%
  # rplot(shape = 15, colours = c("darkorange", "white", "darkcyan")) +
  rplot(colours = c("red", "white", "blue")) +
  theme(axis.text.x = element_text(angle = 90))

corr %>%
  network_plot(min_cor = .3)

# get order
order_vars <- corr %>%
  rearrange() %>%
  .$term

# corr %>%
#   stretch() %>%
#   mutate(x = fct_relevel(x, order_vars)) %>%
#   ggplot(aes(x, abs(r))) +
#  geom_point(alpha = .3, size = 3) +
#    stat_summary(fun = "mean", geom = "pointrange", colour = "black", width = .3) +
#   stat_summary(fun.data = "mean_sdl", fun.args = list(mult = 1), geom = "errorbar", colour = "black", width = .2)
```

Ass correlations: mean only

```{r}
corr <- PI_parameters %>%
  select(-c(sample_id)) %>%
  select(ends_with("mean")) %>%
  correlate() 

corr %>%
  rearrange() %>%
  shave() %>%
  # rplot(shape = 15, colours = c("darkorange", "white", "darkcyan")) +
  rplot(colours = c("red", "white", "blue")) +
  theme(axis.text.x = element_text(angle = 90))

corr %>%
  network_plot(min_cor = .3)

# get order
order_vars <- corr %>%
  rearrange() %>%
  .$term

# corr %>%
#   stretch() %>%
#   mutate(x = fct_relevel(x, order_vars)) %>%
#   ggplot(aes(x, abs(r))) +
#  geom_point(alpha = .3, size = 3) +
#    stat_summary(fun = "mean", geom = "pointrange", colour = "black", width = .3) +
#   stat_summary(fun.data = "mean_sdl", fun.args = list(mult = 1), geom = "errorbar", colour = "black", width = .2)
```

Pairwise correlations

```{r}
PI_parameters %>%
  select(-sample_id) %>%
  select(ends_with("mean")) %>%
  ggplot() +
  ggforce::geom_autopoint() +
  ggforce::geom_autodensity(alpha = .3) +
  ggforce::facet_matrix(vars(everything()), 
                        layer.diag = 2) +
  theme_bw()
```

# Other exploration

consider exploring variation within JE2

```{r}
JE2_parameters <- readRDS("../../Genetic_pairs_analysis_Oct_2020/processed_data/PI/dataframes/parameters/PI_parameters.Rda") %>%
  filter(sample_id == "JE2") %>%
  select(-c(contains("end_point_timepoint"), 
            contains("doubling_time"))) %>%
 select(well_id:end_point_death)

JE2_parameters %>%
  pivot_longer(cols = -well_id, names_to = "parameter", values_to = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~parameter, scales = "free") +
  theme_bw()

scale_recipe <- recipe(~ ., data = JE2_parameters) %>%
  update_role(well_id, new_role = "id") %>%
  step_normalize(all_predictors())

JE2_norm <- prep(scale_recipe) %>%
  bake(new_data = NULL)

JE2_norm %>%
  pivot_longer(cols = -well_id, names_to = "parameter", values_to = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~parameter) +
  theme_bw()

JE2_norm %>%
  pivot_longer(cols = -well_id, names_to = "parameter", values_to = "value") %>%
  ggplot(aes(x = value, fill = parameter)) +
  geom_density(alpha = .5) +
  facet_wrap(~parameter) +
  theme_bw()

# Correlations
corr <- JE2_parameters %>%
  select(-c(well_id)) %>%
  correlate() 

corr %>%
  rearrange() %>%
  shave() %>%
  # rplot(shape = 15, colours = c("darkorange", "white", "darkcyan")) +
  rplot(colours = c("red", "white", "blue")) +
  theme(axis.text.x = element_text(angle = 90))

JE2_parameters %>%
  select(-well_id) %>%
  ggplot() +
  ggforce::geom_autopoint() +
  ggforce::geom_autodensity(alpha = .3) +
  ggforce::facet_matrix(vars(everything()), 
                        layer.diag = 2) +
  theme_bw()
```

# First attempt to perform a PCA

```{r}
pi_recipe <- recipe( ~ ., data = PI_parameters) %>%
  update_role(sample_id, new_role = "id") %>%
  step_normalize(all_predictors())%>%
  step_zv(all_predictors()) 

# pi_recipe <- pi_recipe %>%
#   step_corr(all_predictors())

pi_prep <- prep(pi_recipe)

pi_prep

pi_pca_recipe <- pi_recipe %>%
  step_pca(all_predictors(), num_comp = 5)

pi_pca_estimates <- prep(pi_pca_recipe)

names(pi_pca_estimates)

pi_pca_estimates[[1]]
pi_pca_estimates[[2]]
pi_pca_estimates[[3]]

tidied_pca <- tidy(pi_pca_estimates, 3)

tidied_pca %>%
  filter(component %in% c("PC1", "PC2")) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component) +
  labs(y = NULL)

```

Not sure that this approach is the best. We can simply reproduce what Romain did

```{r}
df_pca <- PI_parameters %>%
  column_to_rownames("sample_id")
pca1 <- dudi.pca(df = df_pca, scannf = F, nf = 5, scale = T)
```

\% variance explained

```{r}
fviz_screeplot(pca1, addlabels = T)
```

```{r}
fviz_contrib(pca1, choice = "var", axes = 1:5) +
  coord_flip()
```

Try with mean only

```{r}
df_pca <- PI_parameters %>%
  select(ends_with("mean"), sample_id) %>%
  select(-contains("doubling")) %>%
  column_to_rownames("sample_id")
pca2 <- dudi.pca(df = df_pca, scannf = F, nf = 5, scale = T)
fviz_screeplot(pca2, addlabels = T)

fviz_contrib(pca2, choice = "var", axes = 1)
fviz_contrib(pca2, choice = "var", axes = 2)
```

# Unsupervised random forest

```{r}
# df_rf <- PI_parameters %>%
#   select(sample_id, ends_with("mean"), ends_with("sd")) 

df_rf <- PI_parameters %>%
  select(sample_id, ends_with("mean"))

rf_preproc <- preProcess(df_rf, method = c("zv"))

df_rf <- predict(rf_preproc, df_rf)

set.seed(1234)

rf1 <- randomForest(x = df_rf[,-1], ntree = 500, proximity = TRUE, importance = TRUE)

rf1

varImpPlot(rf1)

rf1$importanceSD

df_importance <- rf1$importance %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  as_tibble() %>%
  pivot_longer(cols = contains("Mean"), names_to = "importance_type", values_to = "importance") %>%
  mutate(parameter = str_match(Variable, "(.*)_death_(mean|sd)")[,2],
         statistic = str_match(Variable, "(.*)_death_(mean|sd)")[,3]) %>%
  mutate(parameter_long = fct_recode(parameter,
    `PI uptake AUC` = "AUC",
    `Maximum PI uptake rate` = "max_rate",
    `Peak PI uptake` = "max",
    `Time of peak PI uptake` = "time_of_max",
    `Time of maximum PI uptake rate` = "time_of_max_rate",
    `Trough PI uptake`= "min",
    `Time of through PI uptake` = "time_of_min"
  )) %>%
  mutate(parameter_short = fct_recode(parameter,
    `AUC` = "AUC",
    `Maximum rate` = "max_rate",
    `Peak` = "max",
    `Time of peak` = "time_of_max",
    `Time of maximum rate` = "time_of_max_rate",
    `Trough`= "min",
    `Time of through` = "time_of_min",
    `Final uptake` = "end_point"
  )) %>%
  mutate(label = str_c(parameter_short, " (", statistic, ")"))



p1 <- df_importance %>%
  filter(importance_type == "MeanDecreaseGini") %>%
  ggplot(aes(x = fct_reorder(parameter_short, importance), y = importance, fill = statistic)) +
geom_bar(stat = "identity", width = .7, position = "dodge") +
  coord_flip() +
  labs(x = "", y = "Importance") +
  theme_bw()

p2 <- df_importance %>%
  pivot_wider(names_from = "importance_type", values_from = "importance") %>%
  ggplot(aes(x =MeanDecreaseGini, y = MeanDecreaseAccuracy)) +
geom_point(size = 3) +
  geom_text_repel(aes(label = parameter_short)) +
  theme_bw()
  theme(text = element_text(face = "bold"))
```

```{r}
prox <- rf1$proximity
pam.rf <- pam(prox, 3)
pred <- cbind(pam.rf$clustering, PI_parameters$sample_id)
table(pred[,2], pred[,1])
```

```{r}
Clusters <- as.factor(pam.rf$cluster)
Species <- PI_parameters$sample_id
p3 <- PI_parameters %>%
  add_column(cluster = Clusters, .after = "sample_id") %>%
  ggplot(aes(x = AUC_death_mean, 
                            y =log(max_rate_death_mean), 
                            colour = cluster)) + 
  geom_point(size = 3) +
  scale_color_manual(values = c("orange", "red", "blue")) +
  theme_bw()
p4 <- PI_parameters %>%
  add_column(cluster = Clusters, .after = "sample_id") %>%
  ggplot(aes(x = AUC_death_mean, 
                            y =max_rate_death_mean, 
                            colour = cluster)) + 
  geom_point(size = 3) +
  scale_color_manual(values = c("orange", "red", "blue")) +
  theme_bw()
(plot_list <- list(p1, p2, p3,p4))
plot_list %>%
  saveRDS("processed_data/RF_plots.Rda")
```

Plot on PCA axes

```{r}
pca2$li %>%
 filter(Axis1 > 5) 

pca2$li %>%
  ggplot(aes(x = Axis1, y = Axis2, col = Clusters)) +
   scale_color_manual(values = c("orange", "red", "blue")) +
  geom_point() +
  theme_bw()

fviz_pca_biplot(pca2,
             #select.ind = list(contrib = 10),
             axes = c(1,2),
             geom = c("point"),
             alpha.ind = .5,
             addEllipses = F,
             ellipse.level = .95,
             # col.ind = pca_meta.df4$cluster,
             title = paste("" ),
             repel = T,
             pointshape = 16)
```
